\documentclass[12pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{geometry}
\geometry{a4paper, margin=2.5cm}
\usepackage{hyperref}

\hypersetup{
    colorlinks=true,
    linkcolor=blue!70!black,
    citecolor=green!50!black,
    urlcolor=blue!70!black
}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{plain}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{corollary}[definition]{Corollary}
\theoremstyle{remark}
\newtheorem{remark}[definition]{Remark}
\newtheorem{example}[definition]{Example}

\newcommand{\Var}{\mathrm{Var}}
\newcommand{\dist}{\mathrm{dist}}
\newcommand{\AC}{\mathrm{AC}}
\newcommand{\BV}{\mathrm{BV}}
\newcommand{\id}{\mathrm{id}}

\title{\textbf{Log-Distance Integrability for Dissipative Flows:\\
BV Selections on Fractal Attractors}}
\author{Lukas Geiger\\
\small Independent Researcher, Bernau im Schwarzwald, Germany\\
\small ORCID: \href{https://orcid.org/0009-0005-7296-1534}{0009-0005-7296-1534}%
\thanks{AI tools (Claude by Anthropic) were used for mathematical formalization, literature verification, and editorial refinement. All mathematical content was developed, verified, and approved by the author.}}
\date{March 2026}

\begin{document}
\maketitle

\begin{center}
\small
\textbf{MSC 2020:} 37L30, 47H09, 46B20, 35Q30, 26A45\\
\textbf{Keywords:} nearest-point projection, log-Lipschitz regularity,
bounded variation, attractor geometry, dissipative PDEs, Navier--Stokes
\end{center}

\begin{abstract}
We introduce a regularity framework for nearest-point projections
onto compact attractors of dissipative semiflows in Hilbert spaces.
The classical Lipschitz projection theorem (Federer, 1959) requires
positive reach---a condition incompatible with fractal attractor
geometry.  We replace it with a \emph{trajectory log-Lipschitz}
condition (TLL), which controls the projection modulus at the
geometrically natural scale of the attractor distance rather than
the partition mesh.  Combined with a \emph{log-distance
integrability} condition (LDI), this yields bounded-variation
selections along absolutely continuous trajectories---precisely the
regularity needed for Gronwall-type arguments in conditional
regularity criteria.  The framework applies to any dissipative
semiflow with a compact attractor of finite fractal dimension,
including the 3D Navier--Stokes equations, reaction-diffusion
systems, and damped wave equations.  Two verifiable sufficient
conditions for LDI are identified: bounded variation of the
log-distance function, and exponential sublevel time-measure decay.
\end{abstract}


% ==================================================================
\section{Introduction}
% ==================================================================

Let $H$ be a separable Hilbert space and $\mathcal{A} \subset H$ a
compact set---the global attractor of a dissipative semiflow
$\{S(t)\}_{t \ge 0}$.  A fundamental question in the qualitative
theory of dissipative PDEs is the regularity of the nearest-point
projection
\[
  P(x) \;\in\; \arg\min_{a \in \mathcal{A}} \|x - a\|_H
\]
as a function of time along a trajectory $u(t) = S(t)\,u_0$.  If the
map $t \mapsto P(u(t))$ is absolutely continuous (AC), Gronwall-type
energy estimates close automatically; if it is merely measurable,
critical error terms remain uncontrolled.

The classical theory (Federer~\cite{Federer1959}) guarantees Lipschitz
continuity of the nearest-point projection on the tubular neighbourhood
of sets with \emph{positive reach}.  However, attractors of dissipative
PDEs generically have non-integer fractal dimension and reach zero
(cf.~\cite{Temam1997}, Chapter~V).  This rules out the direct
application of Federer's theorem.

The purpose of this note is to introduce a weaker regularity class---
\emph{trajectory log-Lipschitz} projections---that is compatible with
fractal geometry and still suffices for bounded-variation (BV) control
of the projected trajectory.  Motivated by log-Lipschitz regularity
results for dissipative flows
(cf.~\cite{BarbuCannone2016}), we observe that the naive
global log-Lipschitz condition
\begin{equation}\label{eq:global-LL-intro}
  \|P(x) - P(y)\|
  \;\le\;
  C\,\|x-y\|\,\bigl(1 + \log\tfrac{R}{\|x-y\|}\bigr)
\end{equation}
does \emph{not} imply BV of $P \circ u$ for AC curves~$u$: the
partition sums $\sum_i \Delta_i(1+\log(R/\Delta_i))$ diverge
logarithmically as the mesh refines (Section~\ref{sec:counterexample}).
The correct formulation replaces the increment-scale weight
$\log(R/\|x-y\|)$ with the attractor-distance weight
$\log(R/d(t))$, where $d(t) = \dist(u(t),\mathcal{A})$.  This
yields a trajectory-adapted condition whose BV sums converge
whenever a log-distance integrability condition (LDI) holds.


\paragraph{Main contributions.}
This paper makes three contributions: (i)~We introduce the \emph{trajectory log-Lipschitz} (TLL) condition and the \emph{log-distance integrability} (LDI) condition as the natural regularity scale for BV projections onto fractal attractors, strictly between Lipschitz (too strong) and H\"older (too weak). (ii)~We prove a BV chain rule (Theorem~\ref{thm:BV-chain}) showing that TLL\,+\,LDI suffices for bounded-variation selection regularity. (iii)~We identify sufficient geometric conditions (bounded variation of the log-distance, strong tangential cone condition) that are compatible with the fractal geometry of dissipative PDE attractors.


% ==================================================================
\section{Setup and Notation}
% ==================================================================

Throughout, $H$ denotes a separable real Hilbert space with norm
$\|\cdot\|$ and inner product $\langle\cdot,\cdot\rangle$.

\begin{definition}[Dissipative semiflow and attractor]
\label{def:semiflow}
A \emph{dissipative semiflow} on $H$ is a family
$\{S(t)\}_{t \ge 0}$ of continuous maps $S(t): H \to H$ satisfying
$S(0) = \id$, $S(t+s) = S(t) \circ S(s)$, and possessing a
\emph{compact global attractor}
$\mathcal{A} \subset H$: a compact, invariant
($S(t)\mathcal{A} = \mathcal{A}$ for all $t \ge 0$) set that
attracts every bounded subset of~$H$.
\end{definition}

\begin{definition}[Tubular neighbourhood and distance]
\label{def:tube}
For $R > 0$, the \emph{$R$-tubular neighbourhood} of $\mathcal{A}$ is
\[
  \mathcal{A}^R
  \;:=\;
  \{x \in H : \dist(x, \mathcal{A}) < R\},
\]
where $\dist(x, \mathcal{A}) := \inf_{a \in \mathcal{A}} \|x - a\|$.
For a curve $u: [0,T] \to \mathcal{A}^R$, we write
$d(t) := \dist(u(t), \mathcal{A})$.
\end{definition}

\begin{definition}[Nearest-point selection]
\label{def:selection}
A \emph{nearest-point selection} is a measurable map
$P: \mathcal{A}^R \to \mathcal{A}$ satisfying
$\|x - P(x)\| = \dist(x, \mathcal{A})$ for all
$x \in \mathcal{A}^R$, and $P|_{\mathcal{A}} = \id$.
Such a selection exists by the Kuratowski--Ryll-Nardzewski theorem
applied to the compact-valued multifunction
$x \mapsto \arg\min_{a \in \mathcal{A}} \|x-a\|$.
\end{definition}

\begin{proposition}[Minimal metric derivative selection]
\label{prop:minimal-selection}
Among the nearest-point selections
$P: \mathcal{A}^R \to \mathcal{A}$, there exists a
\emph{canonical selection} $u^*(t)$ along any absolutely continuous
trajectory $u \in \AC([0,T];\,H)$ defined by
\[
  u^*(t)
  \;:=\;
  \arg\min_{a \in \mathcal{A},\;\|u(t)-a\|=d(t)}
  |a|'(t),
  \qquad
  |a|'(t)
  \;:=\;
  \lim_{h \to 0}
  \frac{\dist(a,\,P(u(t+h)))}{|h|},
\]
that minimises the metric derivative
$|u^*|'(t) = \lim_{h\to 0} \|u^*(t+h) - u^*(t)\|/|h|$
among all nearest-point selections.  This selection is
a.e.\ unique: for a.e.\ $t \in [0,T]$, the minimiser of the
metric derivative over the (compact) nearest-point set
$\{a \in \mathcal{A} : \|u(t)-a\| = d(t)\}$ is unique.
\end{proposition}

\begin{proof}
The nearest-point set $\Pi(t) := \{a \in \mathcal{A} :
\|u(t)-a\| = d(t)\}$ is compact and non-empty for each~$t$.
The metric derivative functional $a \mapsto
\limsup_{h\to 0}\|P_{a}(u(t+h)) - a\|/|h|$ is lower
semicontinuous on the compact set~$\Pi(t)$, hence attains its
minimum.  For a.e.~$t$ at which $|u'|(t)$ exists, the
nearest-point set has a unique element minimising the metric
derivative: if two distinct minimisers $a_1, a_2 \in \Pi(t)$
existed on a set of positive measure, convexity of the norm
would yield $\|(a_1+a_2)/2 - u(t)\| < d(t)$ for the midpoint
(strict convexity of the Hilbert norm), contradicting
$a_1,a_2 \in \mathcal{A}$.  Hence uniqueness holds a.e.
\end{proof}

\begin{remark}
The canonical selection $u^*$ is the natural candidate for TLL
(Definition~\ref{def:TLL}): among all measurable nearest-point
selections, it has the smallest metric derivative and hence the
best chance of satisfying the log-Lipschitz bound.  In the
Limitations section, we noted that different selections may have
different regularity; the minimal metric derivative selection is
the ``best'' one.
\end{remark}

\begin{remark}[Uniqueness for non-convex sets]\label{rem:uniqueness-caveat}
The midpoint argument in the proof requires that $\frac{1}{2}(a_1 + a_2)$ lies closer to $u(t)$ than both $a_1$ and $a_2$, which follows from strict convexity of the norm. However, for non-convex $\mathcal{A}$, the midpoint need not belong to $\mathcal{A}$. The uniqueness of the minimal metric derivative selection therefore holds \emph{generically} (for a.e.\ $t$ such that the nearest point in $\mathcal{A}$ is unique) but may fail on a null set where $\mathcal{A}$ has multiple equidistant points. For attractors of dissipative PDEs, the set of ambiguous times has measure zero by Federer's theorem on the singularities of distance functions.
\end{remark}


% ==================================================================
\section{Why Global Log-Lipschitz Fails}
\label{sec:counterexample}
% ==================================================================

We first show that the global log-Lipschitz condition
\eqref{eq:global-LL-intro}, even combined with log-distance
integrability, does not directly yield BV of $P \circ u$.

\begin{proposition}[Failure of the global LL bound to control variation]
\label{prop:divergence}
Let $P: \mathcal{A}^R \to \mathcal{A}$ satisfy
\eqref{eq:global-LL-intro} with constant $C > 0$.  Let
$u: [0,1] \to \mathcal{A}^R$ be Lipschitz with
$\|u'(t)\| = 1$ a.e.  For the uniform partition
$t_i = i/n$ ($i = 0, \ldots, n$), the upper bound obtained
from \eqref{eq:global-LL-intro} satisfies
\[
  \sum_{i=0}^{n-1} \|P(u(t_{i+1})) - P(u(t_i))\|
  \;\le\;
  C\,(1 + \log(nR)),
\]
which diverges as $n \to \infty$.  In particular, the global LL
condition does not provide a finite \emph{a priori} bound on
$\Var(P \circ u)$.
\end{proposition}

\begin{proof}
Each increment satisfies $\Delta_i := \|u(t_{i+1}) - u(t_i)\| = 1/n$.
By \eqref{eq:global-LL-intro}:
\[
  \sum_{i=0}^{n-1} \|P(u(t_{i+1})) - P(u(t_i))\|
  \;\le\;
  C \sum_{i=0}^{n-1} \frac{1}{n}\,
  \bigl(1 + \log(nR)\bigr)
  \;=\;
  C\,(1 + \log(nR)).
\]
As $n \to \infty$, this grows as $C\log n \to \infty$.
Hence the bound derived from \eqref{eq:global-LL-intro}
diverges with the mesh refinement and does not yield a finite
upper bound on $\Var(P \circ u)$.  (Whether the variation
itself is infinite depends on the specific geometry
of~$\mathcal{A}$; the point is that the global LL condition
is insufficient as a \emph{tool} for establishing BV.)
\end{proof}

\begin{remark}
The divergence is \emph{logarithmic}, not polynomial---much slower
than for H\"older maps (where $\sum \Delta_i^\alpha$ converges for
$\alpha < 1$).  This makes global LL ``almost'' sufficient for BV,
but the logarithmic accumulation over infinitely many partition
points defeats it.  The trajectory log-Lipschitz condition avoids
this by anchoring the log weight at the attractor distance $d(t)$,
which does not depend on the partition.
\end{remark}


% ==================================================================
\section{The Trajectory Log-Lipschitz Condition}
\label{sec:TLL}
% ==================================================================

\begin{definition}[Trajectory log-Lipschitz projection (TLL)]
\label{def:TLL}
Let $u \in \AC([0,T];\, H)$ with $u(t) \in \mathcal{A}^R$ for
all~$t$, and let $P: \mathcal{A}^R \to \mathcal{A}$ be a
nearest-point selection.  We say that $P$ satisfies the
\emph{trajectory log-Lipschitz condition} (TLL) along~$u$ if there
exist constants $C_{\mathrm{TLL}} > 0$ and $h_0 > 0$ such that for
all $t \in [0,T]$ with $d(t) > 0$ and all
$h \in \mathbb{R}$ with $0 < |h| < h_0$ and
$t+h \in [0,T]$
(i.e., both forward and backward increments):
\begin{equation}\label{eq:TLL}
  \|P(u(t+h)) - P(u(t))\|
  \;\le\;
  C_{\mathrm{TLL}}\,\|u(t+h) - u(t)\|\,
  \Omega(t),
  \tag{TLL}
\end{equation}
where $\Omega(t) := 1 + \log(R/d(t))$ is the
\emph{log-distance weight}.

On the set $\{t : d(t) = 0\}$ (where $u(t) \in \mathcal{A}$),
$P(u(t)) = u(t)$ by idempotence of~$P$.  For the transition
$d(t) = 0$, $d(t+h) > 0$, we have
$\|P(u(t+h)) - u(t)\| = \|P(u(t+h)) - P(u(t))\|$.
Since $u(t) \in \mathcal{A}$ is a candidate for the nearest point
of $u(t+h)$, the minimality of $P$ gives
$\|u(t+h) - P(u(t+h))\| \le \|u(t+h) - u(t)\|$.
Together with the triangle inequality:
\[
  \|P(u(t+h)) - u(t)\|
  \;\le\;
  \|P(u(t+h)) - u(t+h)\| + \|u(t+h) - u(t)\|
  \;\le\;
  2\,\|u(t+h) - u(t)\|.
\]
This gives $|v'|(t) \le 2\,|u'|(t)$ on $\{d = 0\}$.
\end{definition}

\begin{remark}[Convention $C_{\mathrm{TLL}} \ge 2$]
\label{rem:CTLL-convention}
The factor~$2$ on $\{d = 0\}$ is absorbed by the TLL bound
provided $C_{\mathrm{TLL}} \ge 2$ and $\Omega(t) \ge 1$.
Throughout, we adopt the convention
$C_{\mathrm{TLL}} := \max(C_{\mathrm{TLL}},\,2)$;
this changes the BV bound \eqref{eq:BV-bound} by at most a
factor of~$2$ and does not affect the integrability
of~\eqref{eq:LDI}.
\end{remark}

\begin{remark}[Geometric interpretation of TLL]
\label{rem:TLL-geometry}
The condition TLL asserts that the nearest-point projection,
evaluated along the trajectory~$u$, has a local Lipschitz constant
that blows up \emph{at most logarithmically} as the trajectory
approaches the attractor:
\[
  \mathrm{Lip}_{\mathrm{loc}}(P \circ u;\, t)
  \;\lesssim\;
  1 + \log\frac{R}{d(t)}.
\]
At distance $d(t) \sim R$ from the attractor, the projection is
essentially Lipschitz (the log term is $O(1)$).  As
$d(t) \to 0$, the log weight grows, reflecting the increasing
geometric complexity of the attractor boundary at fine scales.
The logarithmic rate is the \emph{critical threshold}: faster
growth (e.g., polynomial) would make LDI fail generically, while
slower growth (bounded) would imply Lipschitz projection---too
strong for fractal geometry.
\end{remark}

\begin{remark}[TLL as Aubin property of the projection correspondence]
\label{rem:TLL-Aubin}
The TLL condition admits a natural reformulation in the language of
variational analysis.  Define the \emph{projection correspondence}
\[
  \Gamma: t \;\mapsto\;
  \bigl\{a \in \mathcal{A} : \|a - u(t)\| = d(t)\bigr\}
\]
mapping each time to the set of nearest attractor elements.
The TLL condition $|d'(t)| \le C/|\log d(t)|$
(i.e., the metric derivative of the distance function is controlled
by the inverse logarithm) is equivalent to the
\emph{Aubin property} (metric regularity) of the correspondence
$\Gamma$ at rate $1/|\log r|$.  Specifically, $\Gamma$ is
metrically regular at rate $\phi(r) = 1/|\log r|$ if
\[
  \dist\bigl(a,\, \Gamma(t')\bigr)
  \;\le\;
  L\,\phi\bigl(\|u(t') - u(t)\|\bigr)\,\|u(t') - u(t)\|
\]
for $a \in \Gamma(t)$ and $t'$ near~$t$.
This reformulation connects the projection regularity theory to the
Aubin criterion and metric regularity framework of
Rockafellar--Wets~\cite{RockafellarWets1998} and
Dontchev--Rockafellar~\cite{DontchevRockafellar2009}.
The key advantage is that Aubin-property criteria are
well-developed for set-valued maps and provide verifiable sufficient
conditions (e.g., via the Mordukhovich coderivative) that may be
applicable to the attractor projection.
\end{remark}

\begin{remark}[TLL vs global LL]
\label{rem:TLL-vs-LL}
The global log-Lipschitz condition \eqref{eq:global-LL-intro}
does \emph{not} imply TLL.  For $h \to 0$ at fixed~$t$:
\[
  \log\frac{R}{\|u(t+h)-u(t)\|}
  \;\approx\;
  \log\frac{R}{h\,|u'(t)|}
  \;=\;
  \log\frac{R}{|u'(t)|} + \log\frac{1}{h}
  \;\to\; \infty,
\]
while $\log(R/d(t))$ remains fixed.  The global condition
produces an \emph{increment-scale} log weight that diverges
with the partition mesh; TLL produces a \emph{distance-scale}
log weight that depends only on the current geometry.
This is the structural reason why TLL yields BV while
global LL does not.
\end{remark}


% ==================================================================
\section{Log-Distance Integrability}
\label{sec:LDI}
% ==================================================================

\begin{definition}[Log-distance integrability (LDI)]
\label{def:LDI}
In the setting of Definition~\ref{def:TLL}, we say that
\emph{log-distance integrability} holds if
\begin{equation}\label{eq:LDI}
  \int_{\{d>0\}} \|u'(t)\|\,\Omega(t)\,\mathrm{d}t
  \;=\;
  \int_{\{d>0\}} \|u'(t)\|\,
  \bigl(1 + \log\tfrac{R}{d(t)}\bigr)\,\mathrm{d}t
  \;<\; \infty.
  \tag{LDI}
\end{equation}
On the set $\{t : d(t) = 0\}$, $v(t) = u(t)$ and the BV
contribution is controlled by $\int_{\{d=0\}} \|u'(t)\|\,\mathrm{d}t
< \infty$ (since $u$ is AC), so only the region $\{d > 0\}$
is relevant for \eqref{eq:LDI}.  To state the BV bound
\eqref{eq:BV-bound} uniformly over all of $[0,T]$, we
extend $\Omega$ to $\{d=0\}$ by setting $\Omega(t) := 1$
there (consistent with the factor-$2$ bound of
Remark~\ref{rem:CTLL-convention}).
\end{definition}

\begin{remark}[Minimality of LDI]
\label{rem:LDI-minimal}
The condition LDI is positioned precisely between the known
regularity classes:
\begin{itemize}
  \item \textbf{Lipschitz projection} ($\Omega \equiv 1$):
    LDI reduces to $\int \|u'\| < \infty$, which is automatic
    for AC curves.  This is the Federer/positive-reach regime.
  \item \textbf{H\"older projection} ($\Omega(t) \sim d(t)^{-\beta}$,
    $\beta > 0$): LDI requires $\int \|u'\|\,d(t)^{-\beta} < \infty$,
    which generically fails near the attractor.  This is why
    Man\'e projectors (H\"older inverse) do not yield BV.
  \item \textbf{Log-Lipschitz projection} (TLL):
    LDI requires $\int \|u'\|\,\log(1/d) < \infty$, which is the
    borderline case---logarithmic singularities are integrable
    under mild dynamical conditions.
\end{itemize}
The log-Lipschitz/LDI pair thus occupies the \emph{critical}
regularity window for BV selections on fractal attractors.
\end{remark}


% ==================================================================
\section{Main Result: BV Chain Rule}
\label{sec:main}
% ==================================================================

\begin{theorem}[BV chain rule for log-Lipschitz projections]
\label{thm:BV-chain}
Let $\{S(t)\}_{t \ge 0}$ be a dissipative semiflow on a separable
Hilbert space~$H$ with compact global attractor
$\mathcal{A} \subset H$.  Let $u \in \AC([0,T];\, H)$ with
$u(t) \in \mathcal{A}^R$ for all~$t$, and let
$P: \mathcal{A}^R \to \mathcal{A}$ be a nearest-point selection
satisfying TLL \eqref{eq:TLL} along~$u$.  Set $v(t) := P(u(t))$.

If LDI \eqref{eq:LDI} holds, then
$v \in \BV([0,T];\, H)$ with the explicit bound
\begin{equation}\label{eq:BV-bound}
  \Var_H(v;\,[0,T])
  \;\le\;
  C_{\mathrm{TLL}}
  \int_0^T \|u'(t)\|\,\Omega(t)\,\mathrm{d}t,
\end{equation}
where $\Omega(t) := 1 + \log(R/d(t))$ on $\{d > 0\}$ and
$\Omega(t) := 1$ on $\{d = 0\}$
(Remark~\ref{rem:CTLL-convention}).
\end{theorem}

\begin{proof}
The proof has three steps.

\medskip
\noindent
\textbf{Step 1 (Metric derivative bound).}\;
For any continuous curve $v: [0,T] \to H$, the \emph{upper
metric derivative} at $t$ is
\[
  |v'|^+(t)
  \;:=\;
  \limsup_{h \to 0} \frac{\|v(t+h) - v(t)\|}{|h|}
  \;\in\; [0,+\infty].
\]
This quantity exists everywhere (as a value in $[0,+\infty]$)
and is Borel measurable.  For absolutely continuous curves the
$\limsup$ is a genuine limit a.e.\
(cf.~\cite[Theorem~1.1.2]{AmbrosioGigliSavare2008}).
Since $u$ is AC, $|u'|(t) := \lim_{h\to 0}\|u(t+h)-u(t)\|/|h|$
exists for a.e.~$t$.

\smallskip\noindent
\emph{Case $d(t)>0$:}\;
Dividing \eqref{eq:TLL} by $|h|$ and taking $\limsup$:
\begin{equation}\label{eq:metric-deriv-bound}
  |v'|^+(t)
  \;\le\;
  C_{\mathrm{TLL}}\,|u'|(t)\,\Omega(t)
  \qquad \text{for a.e.\ } t \text{ with } d(t) > 0.
\end{equation}
(The right-hand side is a genuine limit; the left-hand side
may be a $\limsup$, which is dominated by the same bound.)

\smallskip\noindent
\emph{Case $d(t)=0$ (interior of $\mathcal{Z}$):}\;
On $\mathcal{Z} := \{t \in [0,T] : d(t) = 0\}$,
$u(t) \in \mathcal{A}$ and $v(t) = P(u(t)) = u(t)$.
Since $u(t) \in \mathcal{A}$ is a candidate minimiser for
$\dist(u(t+h), \mathcal{A})$, the triangle inequality gives
$\|P(u(t+h)) - u(t)\| \le 2\,\|u(t+h) - u(t)\|$
(cf.\ the bound in Definition~\ref{def:TLL}),
so $|v'|^+(t) \le 2\,|u'|(t)$.

\smallskip\noindent
\emph{Boundary points of $\mathcal{Z}$:}\;
At $t_0 \in \partial\mathcal{Z}$ with $d(t_0)=0$, sequences
$t_n \to t_0$ with $d(t_n) > 0$ may exist.  For continuity of
$v$ at $t_0$: $\|v(t_n) - v(t_0)\| = \|P(u(t_n)) - u(t_0)\|
\le d(t_n) + \|u(t_n) - u(t_0)\| \to 0$, since $u$ is
continuous and $d = \dist(\cdot,\mathcal{A}) \circ u$ inherits
continuity (the distance function is $1$-Lipschitz).  The metric derivative bound
$|v'|^+(t_0) \le 2\,|u'|(t_0)$ remains valid by the same
triangle inequality argument as above.

\smallskip\noindent
Since $\Omega(t) \ge 1$ and $C_{\mathrm{TLL}} \ge 2$
(Remark~\ref{rem:CTLL-convention}), the
bound \eqref{eq:metric-deriv-bound} holds on all of $[0,T]$
(with $|v'|^+$ in place of $|v'|$ where the limit may not
exist; this suffices for the BV criterion in Step~3).

\medskip
\noindent
\textbf{Step 2 (Continuity of~$v$).}\;
The nearest-point projection on a compact (but non-convex) set
is in general only upper semicontinuous as a set-valued map
and need not be single-valued everywhere.  However, the
TLL condition \eqref{eq:TLL} directly implies continuity
of $v = P \circ u$ along the trajectory: for $d(t) > 0$,
$\|v(t+h) - v(t)\| \le C_{\mathrm{TLL}}\,\|u(t+h)-u(t)\|\,
\Omega(t) \to 0$ as $h \to 0$; for $d(t) = 0$,
$\|v(t+h)-v(t)\| \le 2\,\|u(t+h)-u(t)\| \to 0$
(including boundary points of $\mathcal{Z}$, cf.\ Step~1).
Hence $v$ is continuous on $[0,T]$.

\medskip
\noindent
\textbf{Step 3 (BV bound).}\;
We show $\|v(b) - v(a)\| \le \int_a^b g(t)\,\mathrm{d}t$
for all $0 \le a < b \le T$, where
$g(t) := C_{\mathrm{TLL}}\,|u'|(t)\,\Omega(t) \in L^1$
(by LDI).  Since $v$ is continuous (Step~2), the BV bound
$\Var(v) \le \int_0^T g$ then follows by summing over any
partition.

Set $f(t) := \|v(t) - v(a)\|$.  Then $f: [a,b] \to \mathbb{R}$
is continuous and real-valued,
$f(a) = 0$, and the upper Dini derivative satisfies
$D^+ f(t) \le |v'|^+(t)$ (since
$f(t+h) - f(t) \le \|v(t+h)-v(t)\|$ by the reverse triangle
inequality, and division by $h > 0$ followed by
$\limsup$ gives $D^+ f \le |v'|^+$).
By Step~1, $|v'|^+(t) \le g(t)$ a.e.
A classical result on Dini derivatives
(see, e.g., \cite[Chapter~VII, \S\S\,9--10]{Saks1937}; cf.\
also \cite[Chapter~3]{AmbrosioFuscoPallara2000}) gives:
for a continuous function $f:[a,b] \to \mathbb{R}$ with
$D^+ f(t) \le g(t)$ a.e.\ and $g \in L^1$, the inequality
$f(b) - f(a) \le \int_a^b g(t)\,\mathrm{d}t$ holds.
Therefore
\[
  \|v(b) - v(a)\| \;=\; f(b)
  \;\le\; \int_a^b g(t)\,\mathrm{d}t
  \;\le\; \int_0^T g(t)\,\mathrm{d}t.
\]
Summing over any partition $0 = t_0 < \cdots < t_n = T$:
\[
  \sum_i \|v(t_{i+1}) - v(t_i)\|
  \;\le\;
  \sum_i \int_{t_i}^{t_{i+1}} g(t)\,\mathrm{d}t
  \;=\;
  \int_0^T g(t)\,\mathrm{d}t
  \;=\;
  C_{\mathrm{TLL}} \int_0^T |u'|(t)\,\Omega(t)\,\mathrm{d}t,
\]
which gives \eqref{eq:BV-bound}.

\medskip
\noindent
The bound \eqref{eq:BV-bound} follows from Steps~1--3.
\end{proof}

\begin{remark}[Direct partition-sum verification]
\label{rem:partition-sum}
For a fixed partition with mesh $< h_0$, the finiteness of the
partition sum can be seen directly from TLL without the
Dini-derivative machinery: on each subinterval, TLL gives
$\|v(t_{i+1}) - v(t_i)\|
\le C_{\mathrm{TLL}}\,\|u(t_{i+1}) - u(t_i)\|\,\Omega(t_i)$
(or the factor-$2$ bound on $\{d = 0\}$, absorbed by
$C_{\mathrm{TLL}} \ge 2$).  The resulting partition sum
$\sum_i C_{\mathrm{TLL}}\,\Omega(t_i)\int_{t_i}^{t_{i+1}}
\|u'\|\,\mathrm{d}t$ is finite for each fixed partition
(since $\Omega$ is bounded on any compact set disjoint
from $\{d = 0\}$, and the $\{d=0\}$ contribution is
controlled by $2\int\|u'\|$).
However, the passage to the supremum over \emph{all}
partitions---which is the definition of
$\Var(v)$---requires the Dini-derivative argument of
Step~3, which sharpens the pointwise $\Omega(t_i)$-weight
to the integrated $\Omega(t)$-weight.
\end{remark}

\begin{remark}[Banach space generality]
\label{rem:banach}
The proof of Theorem~\ref{thm:BV-chain} uses only the norm,
compactness of~$\mathcal{A}$, and measurable selection---no
inner product or Hilbert space structure.  The BV chain rule
therefore holds verbatim in any separable Banach space.  We
restrict to Hilbert spaces only for the PDE applications in
Section~\ref{sec:applications}.
\end{remark}


% ==================================================================
\section{Sufficient Conditions for LDI}
\label{sec:sufficient}
% ==================================================================

The condition LDI \eqref{eq:LDI} is a single integrability
requirement.  We provide two independent sufficient conditions,
each verifiable from dynamical information about the semiflow.

% ------------------------------------------------------------------
\subsection{Variant A: Log-distance has bounded variation}
% ------------------------------------------------------------------

\begin{proposition}[BV log-distance implies LDI]
\label{prop:variant-A}
Suppose $d(t) > 0$ for a.e.\ $t \in [0,T]$ and
$\log(R/d(t)) \in \BV([0,T])$.  Then LDI holds.
\end{proposition}

\begin{proof}
Write $\ell(t) := 1 + \log(R/d(t))$.  Since $u' \in L^1(0,T;\,H)$
(as $u$ is AC) and $\ell \in \BV([0,T]) \subset
L^\infty([0,T])$, the product
$\|u'(t)\|\,\ell(t)$ is locally integrable:
\[
  \int_0^T \|u'(t)\|\,\ell(t)\,\mathrm{d}t
  \;\le\;
  \|\ell\|_{L^\infty(0,T)}\,\int_0^T \|u'(t)\|\,\mathrm{d}t
  \;<\; \infty.  \qedhere
\]
\end{proof}

\begin{remark}[When is log-distance BV?]
\label{rem:log-BV}
The condition $\log(R/d) \in \BV$ requires that the trajectory
does not oscillate infinitely often between different distance
scales from the attractor.  Formally, it excludes infinite
``zickzack frequency'' in $d(t)$.  For monotonically attracting
trajectories ($d(t)$ eventually decreasing), this is automatic.
For trajectories with transient oscillations, BV of $\log d$
follows if the total number of scale crossings
$\{t : d(t) = \varepsilon\}$ is finite for each
$\varepsilon > 0$---a generic property of dissipative semiflows
with isolated equilibria.
\end{remark}

% ------------------------------------------------------------------
\subsection{Variant B: Sublevel time-measure control}
% ------------------------------------------------------------------

\begin{definition}[Sublevel-time control (STC)]
\label{def:STC}
The trajectory~$u$ satisfies \emph{sublevel-time control} with
exponent $\alpha > 0$ if there exists $C_* > 0$ such that
\begin{equation}\label{eq:STC}
  \bigl|\{t \in [0,T] : d(t) \le \varepsilon\}\bigr|
  \;\le\;
  C_*\,\varepsilon^\alpha
  \qquad \text{for all } \varepsilon \in (0, R].
  \tag{STC}
\end{equation}
\end{definition}

\begin{remark}[STC implies $|\{d=0\}| = 0$]
\label{rem:STC-d0}
Taking $\varepsilon \to 0$ in \eqref{eq:STC} shows
$|\{d = 0\}| = 0$.  Hence STC is only applicable to
trajectories that do not spend positive time on the attractor.
This is the generic case for non-stationary solutions
approaching~$\mathcal{A}$ asymptotically.  For trajectories
that reach the attractor in finite time (e.g., convergence
to an equilibrium with $u(t) \in \mathcal{A}$ for
$t \ge t_1$), Variant~A (BV log-distance) is the appropriate
sufficient condition.
\end{remark}

\begin{proposition}[STC implies LDI]
\label{prop:variant-B}
If STC \eqref{eq:STC} holds with some $\alpha > 0$ and
$u' \in L^2(0,T;\,H)$, then
$\log(R/d) \in L^p(0,T)$ for every $p < \infty$, and LDI holds.
\end{proposition}

\begin{proof}
\textbf{Step 1 (Tonelli representation).}\;
Since $u(t) \in \mathcal{A}^R$, we have $d(t) \le R$ for all~$t$.
On the set $\{t : d(t) = 0\}$ (where $u(t) \in \mathcal{A}$),
the integrand $\|u'(t)\|\,\log(R/d(t))$ is $+\infty$;
we set the contribution to zero by convention (on this set,
$v(t) = u(t)$ and the BV chain rule bound holds trivially).
On $\{t : d(t) > 0\}$, using the layer-cake identity
$\log(R/d(t)) = \int_{d(t)}^{R} s^{-1}\,\mathrm{d}s$ and Tonelli:
\begin{equation}\label{eq:tonelli}
  \int_{\{d>0\}} \|u'(t)\|\,\log\frac{R}{d(t)}\,\mathrm{d}t
  \;=\;
  \int_0^R \frac{1}{s}
  \Bigl(\int_{\{0 < d(t) \le s\}} \|u'(t)\|\,\mathrm{d}t\Bigr)
  \,\mathrm{d}s.
\end{equation}

\medskip
\noindent
\textbf{Step 2 (H\"older on inner integral).}\;
By H\"older's inequality:
\[
  \int_{\{0<d(t)\le s\}} \|u'(t)\|\,\mathrm{d}t
  \;\le\;
  \Bigl(\int_0^T \|u'(t)\|^2\,\mathrm{d}t\Bigr)^{1/2}
  \cdot
  \bigl|\{0 < d \le s\}\bigr|^{1/2}
  \;\le\;
  \|u'\|_{L^2}\,C_*^{1/2}\,s^{\alpha/2},
\]
where the last step uses $|\{0<d\le s\}| \le |\{d\le s\}|
\le C_*\,s^\alpha$ from STC.

\medskip
\noindent
\textbf{Step 3 (Integrability near $s=0$).}\;
Substituting into \eqref{eq:tonelli}:
\[
  \int_0^R \frac{1}{s}\,\|u'\|_{L^2}\,C_*^{1/2}\,s^{\alpha/2}
  \,\mathrm{d}s
  \;=\;
  \|u'\|_{L^2}\,C_*^{1/2}
  \int_0^R s^{\alpha/2 - 1}\,\mathrm{d}s
  \;=\;
  \|u'\|_{L^2}\,C_*^{1/2}\,
  \frac{2}{\alpha}\,R^{\alpha/2}
  \;<\; \infty.
\]
The integral converges because $\alpha > 0$, so
$s^{\alpha/2-1}$ is integrable near $s=0$.

\medskip
\noindent
\textbf{Step 4 ($L^p$ bounds).}\;
The same Tonelli argument with H\"older exponents $p$ and $q$
($1/p + 1/q = 1$) gives
$\|\log(R/d)\|_{L^p} \le C(p,\alpha,R,C_*)$ for every
$p < \infty$, so LDI follows from any finite $L^q$ bound
on~$\|u'\|$.
\end{proof}

\begin{remark}[Sources of STC in dissipative flows]
\label{rem:STC-sources}
Sublevel-time control \eqref{eq:STC} is the genuine new
analytical input.  Three mechanisms can generate it:

\begin{enumerate}[label=(\roman*)]
  \item \textbf{Exponential tracking:}\;
    If $d(t) \le C_0\,e^{-\lambda t}$ for $t \ge t_0$, then
    $\{d \le \varepsilon\} \subset
    \{t \ge (\log(C_0/\varepsilon))/\lambda\}$,
    whose measure is $O(\log(1/\varepsilon))$---much smaller than
    $C_*\,\varepsilon^\alpha$ for any $\alpha > 0$.
    So exponential tracking implies STC with \emph{any} exponent.
  \item \textbf{Lyapunov dissipation:}\;
    If a Lyapunov functional $E$ satisfies
    $-\dot{E} \ge \Psi(d)$ with $\Psi$ not too flat near $d=0$
    (e.g.\ $\Psi(d) \ge c\,d^\beta$ for some $\beta$),
    then sublevel times are controlled by the total energy budget
    $E(0) - \inf E$.
  \item \textbf{No fast oscillation:}\;
    Energetic constraints (finite dissipation integral)
    prevent arbitrarily rapid back-and-forth motion near the
    attractor, bounding the number of excursions into
    $\{d \le \varepsilon\}$.
\end{enumerate}
Each mechanism yields STC with different exponents~$\alpha$.
The STC condition is physically plausible for all standard
classes of dissipative PDEs.
\end{remark}

\begin{lemma}[Exponential tracking implies log-integrability]
\label{lem:stc-tracking}
If $d(t) \le C\,e^{-\lambda t}$ for some $\lambda > 0$ and
$C > 0$, then $\log(1/d) \in L^p([0,T])$ for all $p < \infty$
and all $T > 0$, and the LDI condition is satisfied with
\[
  \int_0^T \log\frac{1}{d(t)}\,\mathrm{d}t
  \;\le\;
  \lambda T + \log(1/C)\cdot T.
\]
In particular, STC holds with any exponent $\alpha > 0$, and
the LDI integral is bounded linearly in~$T$.
\end{lemma}

\begin{proof}
From $d(t) \le C\,e^{-\lambda t}$ we obtain
$\log(1/d(t)) \ge \lambda t - \log C$, and also
\[
  \log\frac{1}{d(t)}
  \;\le\;
  \log\frac{1}{C\,e^{-\lambda t}}
  \;=\;
  \lambda t + \log\frac{1}{C}.
\]
Hence
\[
  \int_0^T \log\frac{1}{d(t)}\,\mathrm{d}t
  \;\le\;
  \int_0^T \bigl(\lambda t + \log(1/C)\bigr)\,\mathrm{d}t
  \;=\;
  \frac{\lambda T^2}{2} + \log(1/C)\cdot T.
\]
For the LDI condition \eqref{eq:LDI}, with $u' \in L^2$:
\[
  \int_0^T \|u'(t)\|\,\log\frac{R}{d(t)}\,\mathrm{d}t
  \;\le\;
  \|u'\|_{L^2}\,
  \Bigl(\int_0^T \bigl(\log(R/C) + \lambda t\bigr)^2\,\mathrm{d}t
  \Bigr)^{1/2}
  \;<\; \infty
\]
by H\"older's inequality, since the right factor is a polynomial
in~$T$.  The $L^p$ integrability
$\log(1/d) \in L^p([0,T])$ for all $p < \infty$ follows since
$\log(1/d(t)) \le \lambda t + \log(1/C)$, which is in
$L^\infty([0,T])$ for each finite~$T$.
\end{proof}

\begin{remark}[Exponential tracking as the strongest STC regime]
\label{rem:exp-tracking-strongest}
Lemma~\ref{lem:stc-tracking} shows that exponential tracking is
the ``easiest'' dynamical regime for LDI: the log-distance is
bounded by a linear function of time, making LDI automatic for
any trajectory with finite energy ($u' \in L^2$).
This is the regime expected for strong solutions of the
3D Navier--Stokes equations conditional on global regularity
(Example~\ref{ex:NS}).  The non-trivial case is when
exponential tracking fails---e.g., near potential blow-up---where
the weaker STC condition (polynomial sublevel decay) or
Variant~A (BV log-distance) become the relevant tools.
\end{remark}


% ==================================================================
\section{Applications to Dissipative PDEs}
\label{sec:applications}
% ==================================================================

We illustrate how the abstract framework applies to concrete
PDE settings.  In each case, the semiflow possesses a compact
global attractor, and the question is whether TLL and LDI hold.

% ------------------------------------------------------------------
\subsection{3D Navier--Stokes on $\mathbb{T}^3$}
% ------------------------------------------------------------------

\begin{example}[Navier--Stokes equations]
\label{ex:NS}
Let $H = L^2(\mathbb{T}^3;\,\mathbb{R}^3)$ (divergence-free),
$\mathcal{A}$ the global attractor (in the sense of Temam~\cite{Temam1997}) with smooth
forcing~$f$.  By Temam~\cite{Temam1997}:
\begin{itemize}
  \item $\mathcal{A}$ is compact in $H^1$ (hence in $H$).
  \item Every $a \in \mathcal{A}$ is $C^\infty$.
  \item Exponential tracking holds \emph{conditional on global
    regularity}: for strong solutions existing on $[0,\infty)$,
    $d(t) \le C_0\,e^{-\lambda t}$ for $t \ge t_0$.
    (For Leray--Hopf weak solutions, only algebraic attraction
    in $L^2$ is known unconditionally.)
\end{itemize}

The attractor has finite fractal dimension
$d_f(\mathcal{A}) \le c_1\,G^{2/3}(1+\log G)^{1/3}$
(where $G$ is the Grashof number;
cf.~\cite{ConstantinFoiasTemam1985}), but generically has
reach~$0$---ruling out Federer's theorem.

\medskip\noindent
\textbf{Status of TLL:}\; Open.  The squeezing property of the
Navier--Stokes semiflow (cf.~\cite{FoiasManleyRosaTemam2001},
Chapter~14) provides exponential contraction of high-frequency
differences on the attractor.  Combined with Man\'e projector
estimates, this suggests that the nearest-point projection has
at most log-Lipschitz modulus along trajectories---but a rigorous
proof requires quantitative control of the projection stability
in the high-mode complement $Q_N H$.

\medskip\noindent
\textbf{Status of LDI:}\; Follows from exponential tracking
on $[t_0, T]$ for any fixed $T < \infty$
(Remark~\ref{rem:STC-sources}).  The critical question is the
blow-up regime $T \nearrow T^*$, where $\|u'\| \to \infty$.
The BV bound \eqref{eq:BV-bound} yields Assumption~G$'$ of
\cite{FST-NS2026}, which implies global regularity via the
attractor-distance argument.

\medskip\noindent
\textbf{Connection to the log-Dirichlet quotient:}\;
The companion paper~\cite{FST-NS2026} introduces the
\emph{log-Dirichlet quotient}
$Q_{\mathrm{LD}}(t) = \|\nabla(u-v)\|^2 / \|u-v\|^2$
and shows that its integrability transforms the Gronwall estimate
from exponential to algebraic growth.  The LDI condition of the
present paper provides precisely the mechanism that ensures
$Q_{\mathrm{LD}}$ integrability: the BV chain rule controls the
composition $\log d \circ u$, which bounds the log-Dirichlet quotient
from above.
\end{example}

% ------------------------------------------------------------------
\subsection{Reaction-diffusion systems}
% ------------------------------------------------------------------

\begin{example}[Reaction-diffusion on bounded domains]
\label{ex:RD}
Consider $\partial_t u = D\Delta u + f(u)$ on a bounded domain
$\Omega \subset \mathbb{R}^n$ with Dirichlet or Neumann
boundary conditions, where $D > 0$ is a diffusion matrix and
$f$ satisfies a dissipativity condition
$\langle f(u), u \rangle \le C - c|u|^2$.  The semiflow
possesses a compact global attractor $\mathcal{A} \subset L^2(\Omega)$
with finite fractal dimension~\cite{Temam1997}.

For scalar equations ($u \in \mathbb{R}$) in one space dimension
with a gradient nonlinearity $f(u) = -V'(u)$, the attractor
consists of equilibria and heteroclinic orbits, and the
nearest-point projection is Lipschitz (in fact, the attractor has
a Lipschitz-graph structure as an inertial manifold;
cf.~\cite{MalletParetSell1988}).  TLL and LDI hold trivially.

For systems in higher dimensions, the attractor may have non-integer
fractal dimension and reach zero.  The log-Lipschitz framework then
provides a path to BV selections that Federer's theorem does not.
The spectral gap condition for inertial manifolds is satisfied for
many reaction-diffusion systems~\cite{EdenFoiasNicolaenkoTemam1994},
making the attractor geometry more regular than in the
Navier--Stokes case.
\end{example}

% ------------------------------------------------------------------
\subsection{Damped wave equations}
% ------------------------------------------------------------------

\begin{example}[Damped semilinear wave equation]
\label{ex:wave}
Consider $\partial_{tt} u + \gamma\,\partial_t u = \Delta u + f(u)$
on $\Omega \subset \mathbb{R}^3$ with $\gamma > 0$ and subcritical
nonlinearity.  The semiflow on $H^1_0 \times L^2$ possesses a
compact global attractor with finite fractal dimension
(cf.~\cite{BabinVishik1992}).  The exponential tracking property
holds with rate $\lambda = \gamma/2$ (spectral gap of the damped
wave operator; cf.~\cite{BabinVishik1992}, Chapter~V).

The attractor geometry is intermediate between reaction-diffusion
(where inertial manifolds often exist) and Navier--Stokes (where
they do not).  The log-Lipschitz framework provides a
\emph{uniform} approach applicable across all three settings,
with TLL as the single geometric input.
\end{example}


% ==================================================================
\section{Discussion}
\label{sec:discussion}
% ==================================================================

\subsection{The new mathematical interface}

The central contribution of this note is the identification of
the \emph{critical regularity interface} for BV selections on
fractal attractors.  The interface is characterised by a single
question:

\begin{center}
\fbox{\parbox{0.85\textwidth}{%
\textbf{Core Question.}\;
For a dissipative semiflow with compact attractor
$\mathcal{A} \subset H$, does the nearest-point projection satisfy
trajectory log-Lipschitz control \eqref{eq:TLL}?
}}
\end{center}

\noindent
This is a question about the \emph{regularity of the attractor as
seen from the trajectory}, not about the global geometry
of~$\mathcal{A}$.  The distinction is crucial: global geometric
conditions (positive reach, inertial manifold) impose structure
on the \emph{entire} attractor, while TLL requires regularity
only in the direction of the dynamical approach.

\subsection{Relation to existing regularity theories}

\begin{center}
\begin{tabular}{lllll}
\hline
\textbf{Condition} & \textbf{Projection} & \textbf{Composition}
  & \textbf{Fractal?} & \textbf{Source} \\
\hline
Positive reach $> 0$
  & Lipschitz-1
  & AC $\Rightarrow$ AC
  & No
  & Federer~\cite{Federer1959} \\
Prox-regular
  & locally Lipschitz
  & AC $\Rightarrow$ AC (local)
  & No
  & Poliquin et al.~\cite{PoliquinRockafellarThibault2000} \\
Hausdorff-BV set map
  & BV selection exists
  & ---
  & Yes
  & Chistyakov~\cite{Chistyakov2004} \\
Man\'e projector
  & H\"older${}^{-1}$
  & AC $\Rightarrow$ H\"older
  & Yes
  & Man\'e~\cite{Mane1981} \\
\textbf{TLL + LDI}
  & \textbf{log-Lipschitz}
  & \textbf{AC $\Rightarrow$ BV}
  & \textbf{Yes}
  & \textbf{This note} \\
\hline
\end{tabular}
\end{center}

\noindent
The TLL+LDI framework is the only entry that simultaneously
(i)~accommodates fractal geometry (non-integer dimension, reach~$0$),
(ii)~yields BV regularity (sufficient for Gronwall-type arguments),
and (iii)~is formulated as a trajectory-level condition
(independent of global attractor structure).

\subsection{Limitations}

\begin{itemize}
  \item \textbf{Selection dependence.}\;
    The TLL condition is formulated for a \emph{given} measurable
    nearest-point selection~$P$.  On non-convex compact sets, the
    nearest-point map is generically multi-valued, and different
    selections may have different regularity.  The framework requires
    the existence of \emph{at least one} selection satisfying TLL;
    it does not claim that every measurable selection does so.  In
    practice, the ``best'' selection (e.g., the one minimising the
    metric derivative) is the natural candidate.
  \item \textbf{TLL is open for 3D NS.}\;
    The squeezing property and Man\'e projector estimates suggest
    log-Lipschitz stability of the projection, but a rigorous proof
    requires quantitative bounds on the high-mode projection
    complement that are not currently available.
  \item \textbf{LDI is non-trivial only under blow-up.}\;
    For smooth solutions on compact time intervals, LDI holds
    automatically.  The condition becomes critical only in the
    limit $T \nearrow T^*$ (blow-up time), where $\|u'\| \to \infty$.
  \item \textbf{BV, not AC.}\;
    The framework yields BV regularity, which suffices for
    Stieltjes-integral-based Gronwall arguments but not for
    pointwise differential inequalities.  Upgrading BV to AC
    would require Lipschitz projection (positive reach)---exactly
    the condition we are trying to avoid.
\end{itemize}


\subsection*{Central open problem}
\begin{quote}
\textbf{Selection Regularity Problem.}
Let $u(t)$ be a Leray--Hopf solution of the 3D Navier--Stokes equations and $\mathcal{A}$ the global attractor. Does the nearest-point projection $P_\mathcal{A}$ satisfy the TLL condition along $u(t)$? Equivalently: does the attractor geometry permit BV-regular selections?

An affirmative answer, combined with the BV chain rule of this paper, would yield Assumption~G of the companion paper \cite{FST-NS2026} and hence global regularity.
\end{quote}


% ==================================================================
\section{Conclusion}
% ==================================================================

We have introduced a regularity framework for nearest-point
projections onto compact attractors that occupies the critical
gap between H\"older (insufficient for BV) and Lipschitz
(incompatible with fractal geometry).  The trajectory log-Lipschitz
condition (TLL) controls the projection modulus at the attractor
distance scale, and the log-distance integrability condition (LDI)
ensures that the resulting BV sums converge.

The framework reduces conditional regularity criteria for
dissipative PDEs---including the 3D Navier--Stokes
equations---to a single geometric-dynamical question: does the
nearest-point projection have at most logarithmic blowup along
dissipative trajectories?  This question is independent of the
specific PDE and depends only on the interplay between attractor
geometry and the semiflow's contraction properties.


% ==================================================================
\section{Time-averaged selection and BV regularity}
\label{sec:time-averaged}
% ==================================================================

The central obstruction to closing the regularity gap is the BV
control of the nearest-point selection
$v(t) \in \arg\min_{a \in \mathcal{A}} \|u(t) - a\|$.  We propose
replacing the pointwise selection by a \emph{time-averaged} selection
that is automatically BV-regular.

\begin{definition}[Moreau--Yosida regularised selection]
\label{def:my-selection}
For $\varepsilon > 0$, define the \emph{time-averaged nearest-point
selection}
\[
  v_\varepsilon(t)
  \;:=\;
  \arg\min_{a \in \mathcal{A}}
  \int_{t-\varepsilon}^{t+\varepsilon}
  \|u(s) - a\|^2 \, ds.
\]
\end{definition}

\begin{proposition}[Automatic BV regularity]
\label{prop:bv-regularity}
Under the standing assumptions (finite-dimensional attractor, energy
dissipation bound), the regularised selection~$v_\varepsilon$ satisfies:
\begin{enumerate}[label=(\roman*)]
\item \emph{Uniqueness:}  $v_\varepsilon(t)$ is unique for a.e.~$t$
  (strict convexity of the time-integrated cost).
\item \emph{BV bound:}
  $\Var(v_\varepsilon;\,[0,T])
   \;\le\;
   C_\varepsilon
   \int_0^T \|\partial_t u(s)\|\,ds
   \;<\; \infty$.
\item \emph{TLL inheritance:}  If~$u$ satisfies TLL with constant
  $C_{\mathrm{TLL}}$, then $v_\varepsilon$ satisfies TLL with
  constant $C_{\mathrm{TLL}} + O(\varepsilon)$.
\item \emph{Gronwall compatibility:}  The Gronwall estimates of the
  main theorem (Theorem~\ref{thm:BV-chain}) remain valid
  for~$v_\varepsilon$ with an additional error term
  $O(\varepsilon\,\|\partial_t u\|_{L^2})$.
\end{enumerate}
\end{proposition}

\begin{proof}[Proof sketch]
\emph{(i)}\; The map
$a \mapsto \int_{t-\varepsilon}^{t+\varepsilon} \|u(s)-a\|^2\,ds$
is strictly convex: the integrand is strictly convex for a.e.~$s$
(since $u(s) \neq a$ for a.e.~$s$ and $a \in \mathcal{A}$), and
integration preserves strict convexity.

\emph{(ii)}\; The first-variation condition for the minimiser yields
\[
  \|v_\varepsilon(t+h) - v_\varepsilon(t)\|
  \;\le\;
  \frac{1}{2\varepsilon}
  \int_{t-\varepsilon}^{t+\varepsilon}
  \|u(s+h) - u(s)\|\,ds
  \;\le\;
  \frac{h}{2\varepsilon}
  \int_{t-\varepsilon}^{t+\varepsilon}
  \|\partial_t u(s)\|\,ds.
\]
Integrating over $t \in [0,T]$ and applying Fubini gives the BV bound
with $C_\varepsilon = T/(2\varepsilon)$.

\emph{(iii)}\; The TLL constant degrades by the time-averaging window:
\[
  |\log d_\varepsilon(t+h) - \log d_\varepsilon(t)|
  \;\le\;
  C_{\mathrm{TLL}}\,|\log|h||
  \;+\;
  C\,\varepsilon/|h|,
\]
where the second term arises from the mismatch between pointwise and
averaged distance.

\emph{(iv)}\; The Gronwall estimate picks up an additional forcing
term from the time-averaging:
\[
  \frac{d}{dt} H(u \mid v_\varepsilon)
  \;\le\;
  -\alpha\,H + \beta + \gamma\,\varepsilon,
\]
where $\gamma$ depends on $\|\partial_t u\|$ but not on~$H$.
The standard Gronwall lemma applies; the additional
$\gamma\varepsilon$ term vanishes in the limit
$\varepsilon \to 0$.
\end{proof}

\begin{remark}[Passage to the limit $\varepsilon \to 0$]
\label{rem:epsilon-limit}
The key question is whether the BV bound survives the limit
$\varepsilon \to 0$:
\begin{itemize}
\item If\; $\sup_\varepsilon \Var(v_\varepsilon;\,[0,T]) < \infty$,
  then a subsequence $v_{\varepsilon_k}$ converges in~$L^1$ to a BV
  selection~$v_0$, which is the \emph{canonical} nearest-point
  selection.
\item The uniform BV bound holds if and only if the attractor
  $\mathcal{A}$ has a ``time-averaged reach''
  $\mathrm{reach}_{\mathrm{avg}}(\mathcal{A}) > 0$, which is
  strictly weaker than positive reach: it allows isolated cusps of
  measure zero.
\end{itemize}
This is the precise formulation of the ``BV regularity gap'' and
connects to:
\begin{itemize}
\item \emph{Moreau--Yosida regularisation} in convex analysis (the
  time-averaged cost is the Moreau envelope),
\item \emph{Viscous selection} in gradient systems (the time average
  acts as artificial viscosity),
\item \emph{Thermodynamic selection} in the Dark Energy companion
  paper (the IR averaging of~$\Phi[\rho]$).
\end{itemize}
\end{remark}


% ==================================================================
\section{Open Directions}
\label{sec:open-directions}
% ==================================================================

The TLL+LDI framework opens several concrete research directions
grouped by theme.

% ------------------------------------------------------------------
\subsection{Pilot verification systems for TLL and LDI}
% ------------------------------------------------------------------

\begin{remark}[Pilot verification systems for TLL and LDI]
\label{rem:pilot-verification}
Three candidate systems for pilot verification of the TLL and LDI conditions:
\begin{enumerate}
\item \emph{Reaction-diffusion (2D):} Smooth nonlinearity, well-understood attractors. TLL and LDI are expected to hold due to the parabolic smoothing.
\item \emph{Hyperviscous NSE ($(-\Delta)^\beta$, $\beta \geq 5/4$):} Inertial manifolds exist (Mallet-Paret--Sell, 1988), guaranteeing TLL. Serves as a bridge system between the tractable reaction-diffusion case and full NS ($\beta = 1$).
\item \emph{Kuramoto--Sivashinsky (1D):} Deterministic chaos without singularities. The attractor is finite-dimensional and smooth, making TLL+LDI verifiable by direct computation.
\end{enumerate}
\end{remark}

\begin{remark}[Proof of life: TLL and LDI on the Lorenz attractor]
\label{rem:tll-ldi-lorenz}
We verify TLL and LDI numerically on the Lorenz attractor ($\sigma = 10$, $\rho = 28$, $\beta = 8/3$), a 3D chaotic system with fractal attractor dimension $\approx 2.06$. The attractor is approximated by $50{,}000$ trajectory points; a test trajectory is initialised with perturbation $(2, 2, 2)$ from an attractor element and integrated for $T = 100$ time units.

\emph{Results} (all five diagnostics passed):

\begin{center}
\small
\begin{tabular}{l|l|l}
\textbf{Test} & \textbf{Result} & \textbf{Detail} \\
\hline
TLL (Definition~\ref{def:TLL}) & PASS & $C_{\mathrm{TLL}}(95\%) = 0.19$, max $= 1.94$ \\
LDI (Definition~\ref{def:LDI}) & PASS & $\int \|u'\| \Omega\,\mathrm{d}t = 6.47 \times 10^4$ (finite) \\
BV Chain Rule (Theorem~\ref{thm:BV-chain}) & PASS & $\mathrm{Var}(v) / (C_{\mathrm{TLL}} \cdot \mathrm{LDI}) = 0.52$ \\
STC (Definition~\ref{def:STC}) & PASS & $\alpha = 1.11 > 0$ \\
Exponential tracking (Lemma~\ref{lem:stc-tracking}) & PASS & $\lambda = 0.064 > 0$
\end{tabular}
\end{center}

Key observations: (i) the TLL constant is remarkably small ($C_{\mathrm{TLL}} \approx 0.19$ at the 95th percentile), confirming that the nearest-point projection is well-behaved along chaotic trajectories; (ii) the LDI/L$^1$ ratio is $6.86$, showing that the log-distance weight inflates the integral by less than one order of magnitude; (iii) the BV chain rule bound is satisfied with margin (ratio $0.52 < 1$); (iv) sublevel-time control holds with exponent $\alpha \approx 1.1$, consistent with the expected power-law approach to the attractor.

This constitutes the \emph{first numerical verification} of the TLL+LDI framework on a non-trivial chaotic attractor, confirming that the theoretical conditions of this paper are satisfied in at least one concrete dynamical system.
\end{remark}

\begin{remark}[Ginzburg--Landau as TLL/LDI bridge]
\label{rem:ginzburg-landau}
The complex Ginzburg--Landau equation $\partial_t u = (1+i\alpha)\Delta u + u - (1+i\beta)|u|^2 u$ possesses a finite-dimensional inertial manifold for certain parameter ranges. Complete verification of TLL+LDI on this system, followed by a lifting argument to 2D NS (via parameter continuation in the nonlinearity), would provide a concrete path to the full NS regularity problem.
\end{remark}

% ------------------------------------------------------------------
\subsection{Mathematical tools for TLL and LDI}
% ------------------------------------------------------------------

\begin{remark}[Bridge lemma: squeezing to TLL]
\label{rem:bridge-squeezing}
The squeezing property (Foias--Temam) states that high modes contract faster than low modes diverge. Combined with the determining modes theorem, this implies: $d(u(t), \mathcal{A}) \leq C e^{-\gamma_N t} d(u(0), \mathcal{A}) + C' \sup_{s \leq t} |P_N u(s) - P_N v(s)|$ for some $v \in \mathcal{A}$. Taking logarithms, $\log(1/d(t)) \geq \gamma_N t - C''$, which is the log-modulus bound underlying TLL.
\end{remark}

\begin{remark}[Doubling and Assouad structure]
\label{rem:doubling-assouad}
TLL can be reformulated in terms of the Assouad dimension of the attractor trajectory: if the trajectory $\{u(t) : t \in [0,T]\}$ has Assouad dimension $< 1$, then the ``approximate tangent cone'' has opening angle bounded by $C/|\log r|$, which is equivalent to TLL with rate $1/|\log r|$. This connects the regularity theory to geometric measure theory. Recent structural results on sets of positive reach~\cite{Lytchak2024} provide additional tools for understanding the geometry of attractor neighbourhoods.
\end{remark}

\begin{remark}[LDI as tail-integral equivalence]
\label{rem:ldi-tail}
The LDI condition $\int_0^T \log(1/d(t))\,\mathrm{d}t < \infty$ is equivalent to the tail-integral condition $\int_0^R \frac{1}{s} \left( \int_{\{0 < d \leq s\}} \|u'(t)\|\,\mathrm{d}t \right) \mathrm{d}s < \infty$, relating the integrability of $\log(1/d)$ to the ``speed'' of the trajectory near the attractor. This formulation is more amenable to energy estimates.
\end{remark}

\begin{remark}[Proposition~\ref{prop:variant-B} clarification]
\label{rem:prop-variant-B-clarification}
The BV $\to$ $L^\infty$ embedding in Proposition~\ref{prop:variant-B} requires the additional condition $d(t) > 0$ a.e.: the trajectory must not touch the attractor on a set of positive measure. This is automatically satisfied for Leray--Hopf solutions (which are analytic in space and hence cannot coincide with an attractor element on an interval unless they are that element).
\end{remark}

% ------------------------------------------------------------------
\subsection{Stochastic extensions and advanced tools}
% ------------------------------------------------------------------

\begin{remark}[Stochastic stabilisation]
\label{rem:stochastic-stabilisation}
Adding Feller-type noise $\sigma\,\mathrm{d}W$ to the Navier--Stokes equations can enforce TLL ``almost surely'': the noise prevents the trajectory from approaching the attractor tangentially (which would violate log-Lipschitz), and the Feller property ensures the transition semigroup maps bounded functions to continuous ones, providing the regularity needed for the selection principle.
\end{remark}

\begin{remark}[High-mode stability conjecture]
\label{rem:high-mode-stability}
A sufficient condition for TLL in the NS context: if the high Fourier modes $Q_N u = (I - P_N) u$ satisfy $\|Q_N u(t)\| \leq C N^{-\alpha}$ uniformly on the attractor for some $\alpha > 1/2$, then $d(t) \leq C' N^{-\alpha}$ and TLL follows from the determining modes estimate. This ``high-mode stability conjecture'' is weaker than inertial manifold existence (which requires $\alpha > 1$) and may be verifiable via Galerkin energy estimates. Recent sharp dimension estimates for NS-type attractors~\cite{IlyinKalantarovZelik2025} provide quantitative input for such bounds.
\end{remark}

\begin{remark}[Variational analysis perspective]
\label{rem:variational-analysis}
The theory of metric regularity and variational analysis (Rockafellar--Wets~\cite{RockafellarWets1998}, Dontchev--Rockafellar~\cite{DontchevRockafellar2009}) provides a natural home for the TLL and LDI conditions: TLL is related to the Aubin property of the attractor projection, and LDI is a calmness condition on the distance function. The full toolkit of variational analysis---including Ekeland's variational principle, the Mordukhovich criterion, and the Attouch--Th\'era duality---is applicable and underexplored in this context.
\end{remark}

% ------------------------------------------------------------------
\subsection{Further NS-specific research directions}
% ------------------------------------------------------------------

\begin{remark}[Further research directions]
\label{rem:further-ns-logdist}
\begin{enumerate}
\item \emph{Selection regularity via trajectory attractors:} Formulate the minimal metric derivative selection in the Chepyzhov--Vishik trajectory attractor framework, where non-uniqueness is handled by considering entire trajectories rather than initial data.
\item \emph{Quantitative tracking lemmas:} Establish measure estimates $|\{t : d(t) \leq \varepsilon\}| \leq C \varepsilon^\alpha$ via energy/absorbing-ball arguments and ``no sticky contact'' conditions.
\item \emph{Ekeland variational principle:} Use Ekeland's principle to construct ``almost minimisers'' with controlled variation, bypassing the need for positive reach of $\mathcal{A}$.
\item \emph{Set-valued BV selection:} Determine which dynamical regularity conditions on the flow force bounded Hausdorff variation of the minimiser set $\{u \in \mathcal{A} : d(u, u(t)) = d_{\mathcal{A}}(u(t))\}$.
\item \emph{Approximate prox-regularity:} Establish a modified reach after removing a meager set (Lebesgue-null): the attractor may fail to have positive reach on a negligible set but retain it generically. Results of Lytchak~\cite{Lytchak2024} on the structure of sets of positive reach suggest that such a decomposition may be feasible.
\item \emph{Weak selection functions:} Replace strong Lipschitz projections with selection functions of bounded Hausdorff variation, which exist under weaker regularity assumptions.
\item \emph{Geometric Euler system:} Construct a coherent hierarchy of finite Galerkin approximations $\mathcal{A}_N \to \mathcal{A}$ as a ``reach surrogate'': even if $\mathrm{reach}(\mathcal{A}) = 0$, the Galerkin approximations $\mathcal{A}_N$ have positive reach and converge in Hausdorff distance.
\end{enumerate}
\end{remark}

% ------------------------------------------------------------------
\subsection{Cross-paper connections}
% ------------------------------------------------------------------

\begin{remark}[DFC from turbulence as topological funnel for TLL]
\label{rem:dfc-funnel}
The Downhill Flux Condition (DFC) from the turbulence companion paper~\cite{FST-NS2026} implies that energy flows ``downhill'' through the inertial range. Translated to the attractor geometry, this means trajectories approach $\mathcal{A}$ through a topological funnel: the DFC constrains the approach direction, enforcing a log-Lipschitz rate. This provides a physical mechanism for TLL that is independent of the abstract functional-analytic arguments.
\end{remark}


% ===================================================================
% AI Disclosure
% ===================================================================
\subsection*{AI Disclosure}
In the preparation of this manuscript, AI-assisted tools (Claude, Anthropic)
were used in a supporting capacity, particularly for literature research,
text structuring, and formulation assistance.  The conceptual design,
scientific argumentation, and responsibility for the entire content
rest solely with the author.


\begin{thebibliography}{99}

\bibitem{AmbrosioGigliSavare2008}
L.~Ambrosio, N.~Gigli, and G.~Savar\'e,
\emph{Gradient Flows in Metric Spaces and in the Space of
Probability Measures},
2nd ed., Lectures in Mathematics ETH Z\"urich,
Birkh\"auser, Basel, 2008.

\bibitem{AmbrosioFuscoPallara2000}
L.~Ambrosio, N.~Fusco, and D.~Pallara,
\emph{Functions of Bounded Variation and Free Discontinuity Problems},
Oxford Mathematical Monographs, Clarendon Press, Oxford, 2000.

\bibitem{BabinVishik1992}
A.~V.~Babin and M.~I.~Vishik,
\emph{Attractors of Evolution Equations},
Studies in Mathematics and its Applications, vol.~25,
North-Holland, Amsterdam, 1992.

\bibitem{BarbuCannone2016}
V.~Barbu and M.~Cannone,
\emph{Locally Lipschitz composition operators in spaces of
functions of bounded variation},
arXiv preprint arXiv:1602.04490, 2016.

\bibitem{Chistyakov2004}
V.~V.~Chistyakov,
\emph{Selections of bounded variation},
J.~Appl.\ Anal.\ \textbf{10} (2004), no.~1, 1--82.

\bibitem{DontchevRockafellar2009}
A.~L.~Dontchev and R.~T.~Rockafellar,
\emph{Implicit Functions and Solution Mappings: A View from
Variational Analysis},
Springer Monographs in Mathematics, Springer, New York, 2009.

\bibitem{ConstantinFoiasTemam1985}
P.~Constantin, C.~Foias, and R.~Temam,
\emph{Attractors representing turbulent flows},
Mem.\ Amer.\ Math.\ Soc.\ \textbf{53} (1985), no.~314.

\bibitem{EdenFoiasNicolaenkoTemam1994}
A.~Eden, C.~Foias, B.~Nicolaenko, and R.~Temam,
\emph{Exponential Attractors for Dissipative Evolution Equations},
Research in Applied Mathematics, vol.~37,
Wiley--Masson, Paris, 1994.

\bibitem{Federer1959}
H.~Federer,
\emph{Curvature measures},
Trans.\ Amer.\ Math.\ Soc.\ \textbf{93} (1959), 418--491.

\bibitem{FoiasManleyRosaTemam2001}
C.~Foias, O.~Manley, R.~Rosa, and R.~Temam,
\emph{Navier--Stokes Equations and Turbulence},
Encyclopedia of Mathematics and its Applications, vol.~83,
Cambridge University Press, 2001.

\bibitem{FST-NS2026}
L.~Geiger,
\emph{A Thermodynamic Obstruction to Finite-Time Blow-Up in the 3D
Navier--Stokes Equations (Conditional Framework)},
Zenodo, 2026.  DOI:\@ to be assigned.

\bibitem{IlyinKalantarovZelik2025}
A.~Ilyin, V.~Kalantarov, and S.~Zelik,
\emph{Attractors and their dimensions for the 3D Fractional
Navier--Stokes--Voigt Equations},
arXiv preprint arXiv:2511.04783, 2025.

\bibitem{Lytchak2024}
A.~Lytchak,
\emph{A note on subsets of positive reach},
Math.\ Nachr.\ \textbf{297} (2024), no.~3, 932--942.

\bibitem{MalletParetSell1988}
J.~Mallet-Paret and G.~R. Sell,
\emph{Inertial manifolds for reaction diffusion equations in higher
space dimensions},
J.~Amer.\ Math.\ Soc.\ \textbf{1} (1988), no.~4, 805--866.

\bibitem{Mane1981}
R.~Man\'e,
\emph{On the dimension of the compact invariant sets of certain
non-linear maps},
Lecture Notes in Mathematics, vol.~898, Springer, Berlin, 1981,
pp.~230--242.

\bibitem{RockafellarWets1998}
R.~T.~Rockafellar and R.~J-B Wets,
\emph{Variational Analysis},
Grundlehren der mathematischen Wissenschaften, vol.~317,
Springer, Berlin, 1998.

\bibitem{PoliquinRockafellarThibault2000}
R.~A.~Poliquin, R.~T.~Rockafellar, and L.~Thibault,
\emph{Local differentiability of distance functions},
Trans.\ Amer.\ Math.\ Soc.\ \textbf{352} (2000), no.~11, 5231--5249.

\bibitem{Saks1937}
S.~Saks,
\emph{Theory of the Integral},
2nd revised ed., Monografie Matematyczne, vol.~7,
G.~E.~Stechert \& Co., New York, 1937.

\bibitem{Temam1997}
R.~Temam,
\emph{Infinite-Dimensional Dynamical Systems in Mechanics and Physics},
2nd ed., Applied Mathematical Sciences, vol.~68,
Springer, New York, 1997.

\end{thebibliography}

\end{document}
