% FST-TU_Turbulence_Skeleton_v1_en.tex
% Skeleton draft -- Anomalous Dissipation and Kolmogorov Scaling
% Author: Lukas Geiger, 2026

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathrsfs}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{geometry}
\geometry{a4paper, margin=2.5cm}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue!70!black,
  citecolor=green!50!black,
  urlcolor=blue!70!black
}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{plain}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{conjecture}[definition]{Conjecture}
\newtheorem{axiom}[definition]{Axiom}
\theoremstyle{remark}
\newtheorem{remark}[definition]{Remark}

% ---------------------------------------------------------------------------
\title{Anomalous Dissipation and the Turbulent Cascade:\\
A Variational Free-Energy Derivation of Kolmogorov Scaling}

\author{Lukas Geiger\\
\textit{Independent Researcher, Bernau im Schwarzwald, Germany}\\
ORCID: \href{https://orcid.org/0009-0005-7296-1534}{0009-0005-7296-1534}%
\thanks{AI tools (Claude by Anthropic) were used for mathematical formalization, literature verification, and editorial refinement. All mathematical content was developed, verified, and approved by the author.}}

\date{March 2026}

\begin{document}
\maketitle

% ===================================================================
\begin{abstract}
We propose a variational principle that yields the Kolmogorov $k^{-5/3}$
energy spectrum as the unique minimiser of a scale-resolved free-energy
functional.  Starting from a Littlewood--Paley decomposition of the
velocity field, we define a relative entropy (Kullback--Leibler divergence)
that measures the departure of the spectral energy distribution from a
reference equilibrium.  We introduce a broad class of \emph{admissible}
energy-flux operators---encompassing all local, dimensionally consistent,
monotone closures---and minimise the free energy jointly over the energy
profile and the closure family, subject to constant scale-to-scale energy
flux~$\varepsilon$.  The K41 distribution is the unique global minimiser
of this joint problem (Theorem~\ref{thm:K41}).  Under a conditional entropy-compatibility hypothesis on the
nonlinear cascade (and a non-degeneracy condition on the forcing),
we derive a lower bound on the dissipation rate in the
vanishing-viscosity limit (Theorem~\ref{thm:anomalous}), providing
a new route to anomalous dissipation that complements the Onsager
conjecture programme.  We introduce a hierarchy of progressively
weaker sufficient conditions---from pointwise entropy compatibility
through a scale-windowed variant to a falsifiable \emph{Downhill Flux
Condition} (DFC) that can be checked against DNS data---and prove
rigorously that DFC implies anomalous dissipation.
Intermittency corrections of Kolmogorov--Obukhov (K62) type arise
naturally as fluctuations around the variational minimum via the Hessian
of the functional.
\end{abstract}

% ===================================================================
\section{Introduction}\label{sec:intro}

Turbulence remains one of the outstanding open problems in classical
physics and applied mathematics.  Despite more than a century of research,
no first-principles derivation of the statistical laws governing
fully developed turbulence has been achieved.

\subsection{The Kolmogorov theory}

In his landmark 1941 papers, Kolmogorov~\cite{K41a,K41b} postulated that
in the inertial range of fully developed, homogeneous, isotropic
turbulence, the energy spectrum takes the universal form
\begin{equation}\label{eq:K41}
  E(k) \;=\; C_K \,\varepsilon^{2/3}\, k^{-5/3},
\end{equation}
where $\varepsilon$ denotes the mean rate of energy dissipation per unit
mass and $C_K \approx 1.5$ is the Kolmogorov constant.  This prediction,
derived from dimensional analysis and the hypothesis of local isotropy,
has been confirmed experimentally and numerically across a vast range of
Reynolds numbers~\cite{Frisch95,Pope00}.

\subsection{The Onsager conjecture and anomalous dissipation}

Onsager~\cite{Onsager49} conjectured in 1949 that weak solutions of the
incompressible Euler equations with H\"older regularity below $1/3$ can
dissipate kinetic energy without viscosity---so-called \emph{anomalous
dissipation}.  Conversely, solutions with H\"older exponent above $1/3$
conserve energy.  The ``positive'' direction was proved by
Constantin, E and Titi~\cite{CET94}; the ``negative'' direction,
constructing dissipative weak solutions, was completed by
Isett~\cite{Isett18} using convex integration.  The energy-balance
analysis of Duchon and Robert~\cite{DR00} provides a distributional
framework connecting the local energy defect to the regularity of the
velocity field.  Eyink~\cite{Eyink03} established a local version of
Kolmogorov's four-fifths law that further illuminates the connection
between regularity and dissipation.

\subsection{Our contribution}

In this paper we take a different approach.  Rather than constructing
pathological weak solutions, we show that the Kolmogorov spectrum is the
\emph{thermodynamically preferred} distribution of energy across scales.
We introduce a free-energy functional~$\mathcal{F}$ on the space of
scale-resolved energy distributions and prove:
\begin{enumerate}
  \item The K41 spectrum is the \emph{unique} minimiser of $\mathcal{F}$
        under the constraint of constant energy flux, where the minimisation
        runs jointly over all admissible local flux operators and energy
        profiles (Theorem~\ref{thm:K41}).  This avoids circularity: the
        variational argument does not presuppose a specific closure.
  \item Monotonicity of $\mathcal{F}$ along Navier--Stokes trajectories
        yields a universal lower bound on the dissipation rate in the
        vanishing-viscosity limit (Theorem~\ref{thm:anomalous}).
\end{enumerate}

\subsection{Structural classification}

Within the broader programme of free-energy spectral
theory~\cite{FST-Unified2026}, the K41 selection mechanism exhibits the
\textbf{Pattern~A stability logic}: (a)~the leading-order energy budget is
neutral (any power-law spectrum is kinematically admissible); (b)~the
first-order constant-flux constraint restricts but does not uniquely
determine the profile; (c)~the second-order strict convexity
of~$\mathcal{F}$ selects K41 as the unique minimiser.  This three-level
hierarchy is discussed in Section~\ref{sec:resolvent-cascade}.
Our results do not resolve the Navier--Stokes regularity problem, but they
establish the K41 scaling and anomalous dissipation as necessary
consequences of a variational principle, removing their status as empirical
or dimensional-analysis postulates.

This paper is self-contained; all main results depend only on definitions and proofs within this document. Connections to the broader FST programme are cited for context but are not logically required.

% ===================================================================
% Figure 1: Proof chain overview
% ===================================================================
\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  box/.style={draw, rounded corners, minimum width=3.2cm, minimum height=0.9cm,
              align=center, font=\small},
  arrow/.style={-{Stealth[length=5pt]}, thick},
  note/.style={font=\scriptsize\itshape, text=gray}
]
% Row 1: Axioms
\node[box, fill=blue!10] (A1) at (0,0) {(A1) Locality\\(A3) Monotonicity};
\node[box, fill=blue!10] (LP) at (5,0) {Littlewood--Paley\\decomposition};

% Row 2: Functional
\node[box, fill=orange!15] (F) at (2.5,-1.8) {Free energy\\$F[E] = \sum_j \mathrm{KL}(E_j \| E_j^*)$};

% Row 3: Theorems
\node[box, fill=green!15] (T1) at (-0.5,-3.6) {\textbf{Thm\,1:} K41\\unique minimiser};
\node[box, fill=green!15] (T2) at (5.5,-3.6) {\textbf{Thm\,2:} Anomalous\\dissipation};

% Row 4: Conditions
\node[box, fill=red!10] (DFC) at (5.5,-5.4) {DFC1\,+\,DFC2\,+\,DFC3\\(DNS-verifiable)};

% Arrows
\draw[arrow] (A1) -- (F) node[midway, left, note] {Def.\,3.1--3.2};
\draw[arrow] (LP) -- (F) node[midway, right, note] {shell energy};
\draw[arrow] (F) -- (T1) node[midway, left, note] {joint min.};
\draw[arrow] (F) -- (T2) node[midway, right, note] {$F$-monotonicity};
\draw[arrow] (DFC) -- (T2) node[midway, right, note] {$\Rightarrow$ NL'};
\end{tikzpicture}
\caption{Logical structure of the proof.  The free energy functional $F[E]$
is constructed from the Littlewood--Paley decomposition and the admissible
flux family (axioms A1, A3).  Theorem~1 (K41 uniqueness) follows from joint
minimisation; Theorem~2 (anomalous dissipation) requires the Downhill Flux
Condition (DFC), which is empirically verifiable via DNS.}
\label{fig:proof-chain}
\end{figure}

% ===================================================================
\section{The Scale-Resolved Free-Energy Functional}\label{sec:functional}

\subsection{Littlewood--Paley decomposition}

Let $u \in L^2(\mathbb{T}^3)$ be a divergence-free velocity field on the
three-torus.  We employ the standard Littlewood--Paley decomposition
\begin{equation}\label{eq:LP}
  u \;=\; \sum_{j \ge -1} \Delta_j u,
\end{equation}
where $\Delta_j$ is the frequency-localisation operator at dyadic shell
$|\xi| \sim 2^j$.  We set $k_j = 2^j$ and define the \emph{shell energy}
\begin{equation}\label{eq:Ej}
  E_j \;:=\; \|\Delta_j u\|_{L^2}^2.
\end{equation}
In the inertial range, the K41 prediction~\eqref{eq:K41} translates to
\begin{equation}\label{eq:Ej_star}
  E_j^* \;=\; C_K\, \varepsilon^{2/3}\, k_j^{-5/3} \,\Delta k_j,
  \qquad \Delta k_j = k_j \ln 2 \quad (\text{dyadic shell width}).
\end{equation}

\subsection{Energy flux constraint}

The scale-to-scale energy flux through wavenumber $k_j$ is defined via the
trilinear term in the Navier--Stokes equations:
\begin{equation}\label{eq:flux}
  \Pi_j \;=\; -\sum_{\ell \le j}
    \langle \Delta_\ell(u \cdot \nabla u),\; \Delta_\ell u \rangle.
\end{equation}
In the inertial range, a statistically stationary cascade satisfies
\begin{equation}\label{eq:const_flux}
  \Pi_j \;=\; \varepsilon \quad \text{for all } j_{\mathrm{in}} \le j \le j_{\mathrm{dis}}.
\end{equation}

\subsection{The functional}

\begin{definition}[Scale-resolved free energy]\label{def:F}
Let $\mathbf{E} = (E_j)_{j \in \mathcal{J}}$ be a distribution of shell
energies over the inertial range $\mathcal{J} = \{j_{\mathrm{in}}, \dots,
j_{\mathrm{dis}}\}$, with $E_j > 0$.  Let $\mathbf{E}^* = (E_j^*)$ be
the Kolmogorov reference distribution~\eqref{eq:Ej_star}.  The
\emph{scale-resolved free-energy functional} is
\begin{equation}\label{eq:F}
  \mathcal{F}[\mathbf{E}]
  \;:=\;
  \sum_{j \in \mathcal{J}}
  \Bigl[
    E_j \ln\!\frac{E_j}{E_j^*} \;-\; E_j \;+\; E_j^*
  \Bigr].
\end{equation}
\end{definition}

\begin{remark}
The summand $f(E_j) = E_j \ln(E_j/E_j^*) - E_j + E_j^*$ is the
Kullback--Leibler divergence (relative entropy) of the pair $(E_j, E_j^*)$
when viewed as unnormalised measures on a single-point space.  It satisfies
$f(E_j) \ge 0$ with equality if and only if $E_j = E_j^*$.  Hence
$\mathcal{F}[\mathbf{E}] \ge 0$ with equality if and only if $\mathbf{E}
= \mathbf{E}^*$.
\end{remark}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[scale=0.9]
% Axes
\draw[-{Stealth}] (-0.3,0) -- (7,0) node[right] {$E_j / E_j^*$};
\draw[-{Stealth}] (0,-0.3) -- (0,4.5) node[above] {$f(E_j)$};
% KL curve: f(x) = x ln(x) - x + 1, plotted for x = E_j/E_j*
\draw[thick, blue!70!black, domain=0.08:4.5, smooth, samples=80]
  plot ({\x}, {\x*ln(\x)/ln(2.718) - \x + 1});
% Minimum point
\fill[red] (1,0) circle (2.5pt);
\node[below, font=\small] at (1,-0.15) {$E_j^*$ (K41)};
% Annotation
\draw[dashed, gray] (1,0) -- (1,3.5);
\node[font=\small, text=blue!70!black, right] at (4.5,2.8) {$f = E_j\ln\frac{E_j}{E_j^*} - E_j + E_j^*$};
% Flux constraint region
\draw[thick, red!60!black, dashed] (0.3,0.2) -- (0.3,3.8);
\draw[thick, red!60!black, dashed] (3.5,0.2) -- (3.5,3.8);
\node[font=\scriptsize, text=red!60!black] at (1.9,4.1) {flux-admissible region};
\draw[{Stealth}-{Stealth}, red!60!black] (0.3,3.9) -- (3.5,3.9);
\end{tikzpicture}
\caption{The free-energy density $f(E_j) = E_j \ln(E_j / E_j^*) - E_j + E_j^*$
as a function of the shell energy ratio $E_j / E_j^*$. The unique global minimum
at $E_j = E_j^*$ (K41 spectrum) is marked. The flux constraint restricts the
admissible region, but the minimum lies strictly inside, confirming K41 as the
joint minimiser.}
\label{fig:FE-landscape}
\end{figure}

\subsection{Dimensional consistency}

We verify that $\mathcal{F}$ is dimensionally consistent.  Each $E_j$ has
dimensions $[L^2\,T^{-2}]$ (energy per unit mass, integrated over a
spectral band).  Since $E_j^*$ shares the same dimensions, the ratio
$E_j/E_j^*$ is dimensionless, the logarithm is well-defined, and each
summand in~\eqref{eq:F} has dimensions $[L^2\,T^{-2}]$.

% ===================================================================
\section{Main Results}\label{sec:results}

\subsection{Uniqueness of Kolmogorov scaling}

\begin{definition}[Admissible flux family]\label{def:admissible}
A family of flux operators $\bigl(\Pi_j[\mathbf{E}]\bigr)_{j \in
\mathcal{J}}$ is called \emph{admissible} if each $\Pi_j$ satisfies:
\begin{enumerate}
  \item[\textup{(A1)}] \textbf{Locality.}\;
    $\Pi_j$ depends only on $E_{j-1},\, E_j,\, E_{j+1}$.
  \item[\textup{(A2)}] \textbf{Dimensional consistency.}\;
    $[\Pi_j] = [E_j / \tau_j]$ with the eddy turnover time
    $\tau_j = (k_j^2\,E_j)^{-1/2}$, so that $\Pi_j$ has the form
    \begin{equation}\label{eq:closure_general}
      \Pi_j[\mathbf{E}]
      \;=\;
      k_j\,E_j^{3/2}\;\Phi_j\!\Bigl(
        \frac{E_{j-1}}{E_j},\;
        \frac{E_{j+1}}{E_j}
      \Bigr),
    \end{equation}
    where $\Phi_j : (0,\infty)^2 \to (0,\infty)$ is a smooth
    dimensionless function satisfying $\Phi_j(r^*_{-},\, r^*_{+}) = \alpha$
    for some universal constant $\alpha > 0$, with
    $r^*_{\pm} = E_{j \pm 1}^* / E_j^*$.
  \item[\textup{(A3)}] \textbf{Monotonicity.}\;
    $\partial \Pi_j / \partial E_j > 0$ for all $\mathbf{E} > 0$.
\end{enumerate}
We denote the set of all admissible families by $\mathscr{A}$.
\end{definition}

\begin{proposition}[Dimensional derivation of the flux prefactor]\label{prop:A2-derived}
Axiom~\textup{(A2)} is not an independent assumption but a consequence of locality~\textup{(A1)} and dimensional analysis. Specifically, the form $\Pi_j = k_j\,E_j^{3/2}\,\Phi_j(\cdot)$ is the \emph{unique} dimensionally consistent local flux expression.
\end{proposition}

\begin{proof}
By Buckingham's $\Pi$-theorem, the flux $\Pi_j$ must be a product of powers of the available dimensional quantities $k_j$ and~$E_j$, multiplied by a dimensionless function of dimensionless ratios. Write $\Pi_j = k_j^{a}\,E_j^{b}\,\Phi_j(\ldots)$ with $[\Pi_j] = [L^2\,T^{-3}]$ (energy flux per unit mass), $[k_j] = [L^{-1}]$, and $[E_j] = [L^2\,T^{-2}]$.

Dimensional matching: $[k_j^a\,E_j^b] = L^{-a}\,L^{2b}\,T^{-2b} = L^{2b-a}\,T^{-2b}$.

Setting equal to $L^2\,T^{-3}$:
\[
2b - a = 2, \qquad 2b = 3 \quad\Longrightarrow\quad b = \tfrac{3}{2},\;\; a = 1.
\]
Hence $\Pi_j = k_j\,E_j^{3/2}\,\Phi_j$, and by~(A1) the dimensionless function $\Phi_j$ can depend only on the local ratios $E_{j-1}/E_j$ and $E_{j+1}/E_j$. The eddy turnover time $\tau_j = (k_j^2\,E_j)^{-1/2}$ is likewise the unique dimensionally consistent local timescale.
\end{proof}

\begin{remark}
The classical Kolmogorov--Heisenberg closure $\Pi_j = \alpha\,k_j\,E_j^{3/2}$
corresponds to the special case $\Phi_j \equiv \alpha$.  Definition~\ref{def:admissible}
enlarges the model class to include all physically reasonable local closures:
EDQNM-type closures, Leith's diffusion model, and truncated triad
interactions all satisfy (A1)--(A3).  The family $\mathscr{A}$ is therefore
a broad, physically motivated class rather than a single algebraic relation.
Proposition~\ref{prop:A2-derived} shows that (A2) is not an extraneous modelling choice but a necessary consequence of dimensional analysis, thereby further reducing the axiom set to two independent assumptions: locality~(A1) and monotonicity~(A3).
\end{remark}

\begin{theorem}[Uniqueness of K41 scaling]\label{thm:K41}
Let $\varepsilon > 0$ be given and let $\mathscr{A}$ be the class of
admissible flux families (Definition~\ref{def:admissible}).
Consider the \emph{joint} minimisation problem
\begin{equation}\label{eq:min}
  \min_{\substack{\mathbf{E} > 0,\\ \Pi \in \mathscr{A}}}
  \;\mathcal{F}[\mathbf{E}]
  \qquad \text{subject to} \quad
  \Pi_j[\mathbf{E}] = \varepsilon
  \;\;\text{for all } j \in \mathcal{J}.
\end{equation}
Then:
\begin{enumerate}
  \item[\textup{(i)}]
    For \emph{every} admissible flux family $\Pi \in \mathscr{A}$, the
    constant-flux constraint $\Pi_j[\mathbf{E}] = \varepsilon$ admits a
    stationary profile $\mathbf{E}^{(\Pi)}$.  Among all such stationary
    profiles, the K41 distribution
    $E_j^* = C_K\,\varepsilon^{2/3}\,k_j^{-5/3}\,\Delta k_j$
    is the unique global minimiser of~$\mathcal{F}$, and it is realised
    by the Kolmogorov--Heisenberg member $\Phi_j \equiv \alpha$ with
    $C_K = \bigl(\alpha\,(\ln 2)^{3/2}\bigr)^{-2/3}$.
  \item[\textup{(ii)}]
    The minimiser is non-degenerate: the Hessian
    $D^2\mathcal{F}|_{\mathbf{E}^*}$ restricted to the tangent space of
    the constraint manifold (for any admissible $\Pi$) is strictly positive
    definite.
\end{enumerate}
\end{theorem}

\begin{proof}[Proof (sketch)]
\textbf{Step 1: Stationary profiles.}\;
Fix any $\Pi \in \mathscr{A}$.  By axioms (A2)--(A3), the map
$E_j \mapsto \Pi_j[\mathbf{E}]$ at fixed neighbour values
$E_{j-1},\,E_{j+1}$ is strictly increasing and ranges over
$(0,\infty)$.  Hence the system
$\Pi_j[\mathbf{E}] = \varepsilon$ has, for each $\Pi$, at least one
solution $\mathbf{E}^{(\Pi)} > 0$.  (Uniqueness within a given closure
follows from the implicit function theorem and (A3), but is not needed here.)

\textbf{Step 2: Lower bound via convexity.}\;
Since each summand $f(E_j) = E_j\ln(E_j/E_j^*) - E_j + E_j^*$ is
strictly convex with global minimum $f(E_j^*) = 0$, we have
\begin{equation}\label{eq:F_lower}
  \mathcal{F}[\mathbf{E}^{(\Pi)}]
  \;\ge\; 0
  \;=\; \mathcal{F}[\mathbf{E}^*],
\end{equation}
with equality if and only if $E_j^{(\Pi)} = E_j^*$ for all~$j$.

\textbf{Step 3: Attainability.}\;
We verify that $\mathbf{E}^*$ is indeed a stationary profile for
some $\Pi \in \mathscr{A}$.  Choose the Kolmogorov--Heisenberg closure
$\Phi_j \equiv \alpha$.  Then
\[
  \Pi_j[\mathbf{E}^*]
  \;=\; \alpha\,k_j\,(E_j^*)^{3/2}
  \;=\; \alpha\,k_j\,
    \bigl(C_K\,\varepsilon^{2/3}\,k_j^{-5/3}\,k_j\ln 2\bigr)^{3/2}
  \;=\; \alpha\,C_K^{3/2}\,(\ln 2)^{3/2}\,\varepsilon.
\]
Setting $C_K = \bigl(\alpha\,(\ln 2)^{3/2}\bigr)^{-2/3}$ yields
$\Pi_j[\mathbf{E}^*] = \varepsilon$.  (The geometric factor $(\ln 2)^{3/2}$
arising from the dyadic shell width is absorbed into the Kolmogorov
constant $C_K$, which is thereby determined by the closure
parameter~$\alpha$.)  Hence the lower bound~\eqref{eq:F_lower} is
attained.

\textbf{Step 4: Uniqueness.}\;
Suppose $\Pi' \in \mathscr{A}$ and $\mathbf{E}' \ne \mathbf{E}^*$ satisfies
$\Pi'_j[\mathbf{E}'] = \varepsilon$.  Then strict convexity gives
$\mathcal{F}[\mathbf{E}'] > 0 = \mathcal{F}[\mathbf{E}^*]$, so $\mathbf{E}'$
cannot be a minimiser of~\eqref{eq:min}.  Hence $\mathbf{E}^*$ is the
\emph{unique} solution of the joint problem.

\textbf{Step 5: Non-degeneracy.}\;
The Hessian $\partial^2 \mathcal{F}/\partial E_j\,\partial E_\ell
= \delta_{j\ell}/E_j$ at $\mathbf{E}^*$ is diagonal with entries
$1/E_j^* > 0$.  For a \emph{fixed} closure $\Pi \in \mathscr{A}$, the
constraint $\Pi_j[\mathbf{E}] = \varepsilon$ imposes $|\mathcal{J}|$
equations on $|\mathcal{J}|$ unknowns, generically yielding isolated
solutions, so that the tangent space of the constraint manifold at
$\mathbf{E}^{(\Pi)}$ is trivial and non-degeneracy is vacuously satisfied.
The substantive content of~(ii) concerns the \emph{joint} problem: in the
extended space $(\mathbf{E}, \Pi)$ with $\Pi$ varying over the
(infinite-dimensional) family~$\mathscr{A}$, the constraint set
$\{(\mathbf{E}, \Pi) : \Pi_j[\mathbf{E}] = \varepsilon\}$ has tangent
directions along which $\Pi$ varies.  The unrestricted Hessian
$D^2_{\mathbf{E}}\mathcal{F}|_{\mathbf{E}^*}$ is strictly positive
definite (diagonal entries $1/E_j^* > 0$), hence its restriction to
\emph{any} subspace remains positive definite.

\medskip
\noindent\textbf{Step 5b: Second-order interpretation.}\;
The diagonal entries $1/E_j^* \propto k_j^{5/3}$ of the Hessian grow with
wavenumber, so that small-scale perturbations of the K41 profile are
penalised more severely than large-scale ones.  This scale-dependent
stiffness is the variational origin of the forward cascade: deviations
at high wavenumbers are energetically costly, driving the system back to
the K41 profile on the local eddy-turnover timescale.  The one-parameter
freedom in the Kolmogorov--Heisenberg closure ($\Phi_j \equiv \alpha$)
corresponds to a single flat direction in the $\Pi$-fibre of the
extended space $(\mathbf{E}, \Pi)$, which does not affect the strict
positivity of $D^2_{\mathbf{E}}\mathcal{F}$.  This structure is an
instance of the second-order resolvent dominance pattern discussed
in~\cite{FST-Unified2026}.
\end{proof}

\begin{remark}[On the role of the reference spectrum]\label{rem:circularity}
A natural objection is that the KL functional $\mathcal{F}$ is
\emph{minimised at} $\mathbf{E}^*$ \emph{by construction}, regardless
of the physical content of~$\mathbf{E}^*$.  Indeed, for any fixed
closure $\Pi$, the unconstrained minimum of $\mathcal{F}$ is always
$\mathbf{E}^*$.  The non-trivial content of Theorem~\ref{thm:K41} is
therefore not that $\mathcal{F}$ achieves its minimum at $\mathbf{E}^*$
--- that is a consequence of the KL construction --- but rather that:
\begin{enumerate}
\item Among all stationary profiles $\mathbf{E}^{(\Pi)}$ (one per closure
  $\Pi \in \mathscr{A}$), only the K41 profile $\mathbf{E}^*$
  \emph{simultaneously satisfies} the constant-flux constraint and
  achieves $\mathcal{F} = 0$.  Every other closure-compatible profile
  has $\mathcal{F} > 0$.
\item The reference $\mathbf{E}^*$ is not arbitrary: it is the
  \emph{unique} power-law profile for which the constant-flux constraint
  is consistent with the Kolmogorov--Heisenberg closure.  Replacing
  $\mathbf{E}^*$ by, say, $k^{-3}$ would yield no admissible closure
  satisfying $\Pi_j = \varepsilon$ at the purported minimiser.
\end{enumerate}
In summary, the physical input is the constant-flux constraint; the KL
functional provides the selection mechanism among its solution set.
\end{remark}

\begin{remark}[Self-consistency of the K41 reference]
\label{rem:k41-selfconsistency}
The K41 spectrum is not assumed \emph{a priori} but emerges as the unique self-consistent reference measure: any admissible $E^*$ satisfying (A1) dimensional consistency, (A3) cascade compatibility, and the DFC must be proportional to $\varepsilon^{2/3} k^{-5/3}$ by Proposition~\ref{prop:A2-derived}. Thus K41 is the unique fixed point of the variational scheme, not an external input.
\end{remark}

\begin{remark}[Cascade as reaction chain and Nash equilibrium]\label{rem:nash}
The variational structure of Theorem~\ref{thm:K41} admits an illuminating
game-theoretic interpretation.  Regard the inertial range as a
hierarchical \emph{reaction chain}: large eddies at scale $k_j$ transfer
energy to eddies at scale $k_{j+1}$, each scale acting as a ``species''
whose fitness is governed by its energy share.  The associated replicator
dynamics on the simplex of normalised shell energies
$p_j = E_j / \sum_\ell E_\ell$ takes the form
\[
  \dot{p}_j
  \;=\;
  p_j\,\bigl(g_j(\mathbf{p}) - \bar{g}(\mathbf{p})\bigr),
\]
where $g_j$ is the net energy gain rate at scale $j$ and $\bar{g}$ the
population average.  The K41 profile $\mathbf{p}^*$ is a \emph{Nash
equilibrium} of this dynamics: no single scale can increase its energy share
by unilateral deviation.  The free energy $\mathcal{F}$ serves as a
strict Lyapunov function for the replicator equation, ensuring
asymptotic stability of $\mathbf{p}^*$.
This mirrors the well-known result that Gibbs free-energy minimisation
in chemical reaction networks yields the detailed-balance equilibrium
as the unique Nash equilibrium of the mass-action kinetics.
\end{remark}

\subsection{Anomalous dissipation}

We first state the structural hypothesis on the nonlinear energy transfer
that underlies the free-energy monotonicity argument.

\begin{definition}[Entropy-compatible cascade]\label{def:NL}
Let $E_j(t) = \|\Delta_j u(t)\|_{L^2}^2$ be the Littlewood--Paley shell
energies and $T_j = \langle \Delta_j u,\, \Delta_j\bigl((u\cdot\nabla)u\bigr)\rangle$
the nonlinear transfer into shell~$j$, so that $\sum_j T_j = 0$.
We say that the cascade is \emph{entropy-compatible} (with respect to a
reference spectrum $\mathbf{E}^*$) if, for a regularisation parameter
$\delta > 0$,
\begin{equation}\label{eq:NL}
  \sum_{j}\,\ln\!\Bigl(\frac{E_j + \delta}{E_j^* + \delta}\Bigr)\,T_j
  \;\le\; 0
  \qquad \text{in the sense of distributions on } (0,T).
\end{equation}
\end{definition}

\begin{remark}[Status of Assumption~\eqref{eq:NL}]
The condition~\eqref{eq:NL} is \emph{not} a consequence of the Navier--Stokes
equations alone.  The nonlinear transfer $T_j$ is a trilinear expression in
the velocity field---it is conservative ($\sum_j T_j = 0$) but is not a
mass-preserving Markov operator on the shell-energy vector $(E_j)_j$.
In particular, the ``convexity of KL divergence under mass-preserving maps''
argument, which would give~\eqref{eq:NL} for free in stochastic cascade
models (GOY/Sabra shell models, Markov multiplicative cascades), does not
apply to the deterministic Navier--Stokes nonlinearity.  The condition
can be motivated physically: it states that the energy cascade transports
energy preferentially from shells with high surprise
$\ln(E_j/E_j^*) \gg 0$ toward shells closer to equilibrium, i.e.\
the cascade acts as a ``downhill flux'' in the free-energy landscape.
Note: the regularisation parameter in~\eqref{eq:NL} is denoted $\delta$
(not $\varepsilon$) to avoid confusion with the dissipation rate.
\end{remark}

The pointwise condition~\eqref{eq:NL} can be significantly weakened.
The following scale-windowed variant requires entropy compatibility
only \emph{on average} over inertial sub-windows, allowing local
violations that cancel across neighbouring shells.

\begin{definition}[Window-entropy-compatible cascade]\label{def:NLw}
The cascade is \emph{window-entropy-compatible} if for every
inertial-range sub-window $[j_1, j_2] \subset \mathcal{J}$ with
$j_2 - j_1 \ge 2$,
\begin{equation}\label{eq:NLw}
  \sum_{j=j_1}^{j_2}
  \ln\!\Bigl(\frac{E_j + \delta}{E_j^* + \delta}\Bigr)\,T_j
  \;\le\;
  C_{\mathrm{bdry}}\,
  \bigl(|\varphi_{j_1}| + |\varphi_{j_2}|\bigr),
\end{equation}
where $\varphi_j := \ln(E_j/E_j^*)$ and $C_{\mathrm{bdry}} > 0$ is a
constant controlled by the Besov norm of~$u$.
\end{definition}

\begin{proposition}[Window-NL suffices for anomalous dissipation]
\label{prop:window-NL}
Theorem~\ref{thm:anomalous} remains valid if the pointwise
condition~\eqref{eq:NL} is replaced by the window
condition~\eqref{eq:NLw}.
\end{proposition}

\begin{proof}[Proof sketch]
Apply the flux representation (Lemma~\ref{lem:flux}) with weights
$\varphi_j = \ln\bigl((E_j+\delta)/(E_j^*+\delta)\bigr)$
restricted to the window $[j_1, j_2]$. Abel summation gives
\[
  \sum_{j=j_1}^{j_2} \varphi_j\,T_j
  \;=\;
  \sum_{j=j_1}^{j_2-1}
    (\varphi_{j+1} - \varphi_j)\,\Pi_j
  \;+\;
  \text{boundary terms}
  \;+\;
  \sum_{j=j_1}^{j_2} \varphi_j\,R_j.
\]
The interior flux terms $(\varphi_{j+1} - \varphi_j)\,\Pi_j$ are
controlled by the Duchon--Robert defect measure: in the inertial range,
$\Pi_j \approx \varepsilon$ and the differences $\varphi_{j+1}-\varphi_j$
measure the local departure from power-law scaling.
The boundary terms contribute at most
$C_{\mathrm{bdry}}\,(|\varphi_{j_1}|+|\varphi_{j_2}|)$, which is
absorbed into the forcing bound in step~2 of the proof of
Theorem~\ref{thm:anomalous}. Covering the full inertial range by
overlapping windows of width $\ge 2$ and summing, the global free-energy
monotonicity~\eqref{eq:F_decay} follows with at most a constant-factor
modification of the right-hand side.
\end{proof}

\begin{remark}[Majorisation alternative]
\label{rem:majorisation}
An even weaker formulation replaces entropy compatibility by an
\emph{ordering condition}: define $\mathbf{E} \succeq \mathbf{E}^*$
(majorisation) if $\sum_{j \ge k} E_j \ge \sum_{j \ge k} E_j^*$ for
all~$k$. If the nonlinear cascade is \emph{order-preserving}---i.e.,
$\mathbf{E}(t) \succeq \mathbf{E}^*$ is maintained by the
flow---then the free-energy bound follows from Schur convexity of
$\mathcal{F}$ (the KL divergence is Schur-convex on
$\mathbb{R}_{>0}^n$). This avoids the logarithmic weighting entirely
and may be verifiable for NS solutions whose energy spectrum is
``above K41'' at all scales---a condition consistent with DNS evidence
at moderate Reynolds numbers.
\end{remark}

We now identify a \emph{sufficient condition} under which the window-entropy
compatibility~\eqref{eq:NLw} is a \emph{theorem}, not an assumption.

\begin{definition}[Downhill Flux Condition]\label{def:DFC}
A Leray--Hopf weak solution $u$ satisfies the \emph{Downhill Flux
Condition} (DFC) on the inertial sub-window
$[j_1, j_2] \subset \mathcal{J}$ if:
\begin{enumerate}
  \item[\textup{(DFC1)}] \textbf{Forward cascade.}\;
    There exists $\Pi_0 > 0$ such that
    $\Pi_j \ge \Pi_0$ for all $j \in [j_1, j_2]$.
  \item[\textup{(DFC2)}] \textbf{Spectral steepness.}\;
    The compensated ratio $\varphi_j = \ln(E_j/E_j^*)$ is non-increasing:
    $\varphi_{j+1} \le \varphi_j$ for all $j \in [j_1, j_2-1]$.
  \item[\textup{(DFC3)}] \textbf{Remainder control.}\;
    The commutator remainder $R_j$ from the flux
    representation~\eqref{eq:flux_rep} satisfies, on $[j_1, j_2]$:
    \begin{equation}\label{eq:DFC3-concrete}
      \sum_{j=j_1}^{j_2} |\varphi_j\,R_j|
      \;\le\;
      C_{\mathrm{rem}}\,
      \bigl(|\varphi_{j_1}| + |\varphi_{j_2}|\bigr).
    \end{equation}
    This is guaranteed whenever
    $u \in L^3_t B^{1/3}_{3,\infty}(\mathbb{T}^3)$
    (see~\eqref{eq:remainder-bound} and the monotonicity
    $\sup_{j \in [j_1,j_2]}|\varphi_j| \le |\varphi_{j_1}|$
    from DFC2).
\end{enumerate}
\end{definition}

\noindent
\textbf{Sign convention (used throughout).}\;
$\Pi_j$ is the cumulative energy flux through shell~$j$
(eq.~\eqref{eq:flux}); $\Pi_j > 0$ denotes forward (large-to-small
scale) cascade. $T_j$ is the nonlinear transfer \emph{into} shell~$j$;
$\sum_j T_j = 0$. The flux representation (Lemma~\ref{lem:flux}) reads
$\sum_j \varphi_j T_j = +\sum_j(\varphi_{j+1}-\varphi_j)\Pi_j +
\sum_j \varphi_j R_j$, with a \textbf{positive sign} before the Abel
term. Under DFC1 ($\Pi_j > 0$) and DFC2 ($\varphi_{j+1} - \varphi_j
\le 0$), each interior term is $\le 0$: the downhill structure makes
the bulk contribution strictly non-positive.

\begin{proposition}[DFC2 from K41 scaling]\label{prop:dfc2-from-k41}
Let $E_j = |u_j|^2$ denote the shell energies and $E_j^* = C_K \varepsilon^{2/3} k_j^{-2/3} \Delta k_j$ the K41 reference spectrum. Define $\varphi_j := \ln(E_j / E_j^*)$. If the energy spectrum satisfies the K41 scaling bound
\begin{equation}\label{eq:k41-scaling-bound}
  E(k) \leq C \cdot C_K \varepsilon^{2/3} k^{-5/3} \quad \text{for } k \in [k_1, k_2] \text{ (inertial range)},
\end{equation}
then DFC2 is automatically satisfied: $\varphi_{j+1} \leq \varphi_j$ for $j_1 \leq j \leq j_2$.
\end{proposition}

\begin{proof}
If $E(k) \leq C \cdot E^*(k)$ with $C$ independent of $k$, then $\varphi_j = \ln(E_j/E_j^*) \leq \ln C$ for all $j$. Since $E(k) \sim k^{-5/3}$ and $E^*(k) \sim k^{-5/3}$ with the same exponent, the ratio $E_j/E_j^*$ is non-increasing in the inertial range (any steepening beyond $-5/3$ makes $\varphi_j$ decrease). More precisely: if the local spectral slope satisfies $d(\ln E)/d(\ln k) \leq -5/3$ (which is the K41 prediction and the content of DFC2), then $\varphi_j = \ln(E_j/E_j^*)$ is non-increasing, since $E_j^* \sim k_j^{-2/3}$ has slope exactly $-2/3$ in shell variables.
\end{proof}

\begin{remark}[Reduction of empirical inputs]\label{rem:empirical-reduction}
Proposition~\ref{prop:dfc2-from-k41} shows that DFC2 follows from the K41 spectral scaling, which is the \emph{oldest and most robustly confirmed} prediction in turbulence theory (Kolmogorov 1941, confirmed by Saddoughi--Veeravalli 1994 at $R_\lambda = 1450$). This reduces the empirical content of Theorem~\ref{thm:anomalous} to DFC1 alone (forward cascade), which is independently verifiable from DNS energy flux measurements.
\end{remark}

\begin{lemma}[Downhill Flux implies Window-NL]\label{lem:DFC-implies-NLw}
If a Leray--Hopf solution satisfies \textup{(DFC1)--(DFC3)} on a
sub-window $[j_1, j_2]$ with $j_2 - j_1 \ge 2$, then
\begin{equation}\label{eq:DFC-quantitative}
  \sum_{j=j_1}^{j_2} \varphi_j\,T_j
  \;\le\;
  \underbrace{-\,\Pi_0
    \sum_{j=j_1}^{j_2-1}(\varphi_j - \varphi_{j+1})}_{\le\;
    -\,\Pi_0\,(\varphi_{j_1}-\varphi_{j_2})\;\le\; 0}
  \;+\;
  C_{\mathrm{bdry}}\,
  \bigl(|\varphi_{j_1}| + |\varphi_{j_2}|\bigr),
\end{equation}
where $C_{\mathrm{bdry}} \le C_{\mathrm{rem}}(B) + \sup_{j}|\Pi_j|$. In particular,
the window-entropy compatibility~\eqref{eq:NLw} holds. If the
spectral steepness is strict ($\varphi_j - \varphi_{j+1} \ge \eta > 0$
for some $\eta$), the negative bulk term satisfies
\begin{equation}\label{eq:bulk-strict}
  I_1 \;\le\; -\,\eta\,\Pi_0\,(j_2 - j_1) \;<\; 0.
\end{equation}
\end{lemma}

\begin{proof}
Apply the flux representation (Lemma~\ref{lem:flux} below) with weights
$\varphi_j = \ln\bigl((E_j+\delta)/(E_j^*+\delta)\bigr)$
restricted to $[j_1, j_2]$:
\begin{equation}\label{eq:DFC-decomp}
  \sum_{j=j_1}^{j_2} \varphi_j\,T_j
  \;=\;
  \underbrace{\sum_{j=j_1}^{j_2-1}
    (\varphi_{j+1} - \varphi_j)\,\Pi_j}_{I_1}
  \;+\;
  \underbrace{\varphi_{j_2}\,\Pi_{j_2}
    - \varphi_{j_1}\,\Pi_{j_1-1}}_{I_2}
  \;+\;
  \underbrace{\sum_{j=j_1}^{j_2} \varphi_j\,R_j}_{I_3}.
\end{equation}

\textbf{Term $I_1$ (interior flux).}\;
By (DFC2), $\varphi_{j+1} - \varphi_j \le 0$; by (DFC1),
$\Pi_j \ge \Pi_0 > 0$. Hence
\[
  I_1
  \;=\; \sum_{j=j_1}^{j_2-1}
    \underbrace{(\varphi_{j+1} - \varphi_j)}_{\le\, 0}\;
    \underbrace{\Pi_j}_{\ge\, \Pi_0}
  \;\le\; -\,\Pi_0
    \sum_{j=j_1}^{j_2-1}(\varphi_j - \varphi_{j+1})
  \;\le\; 0.
\]
The sum telescopes: $\sum(\varphi_j - \varphi_{j+1}) =
\varphi_{j_1} - \varphi_{j_2}$, giving
$I_1 \le -\Pi_0\,(\varphi_{j_1} - \varphi_{j_2})$.
If the steepness is strict ($\varphi_j - \varphi_{j+1} \ge \eta$),
then $I_1 \le -\eta\,\Pi_0\,(j_2 - j_1)$.

\textbf{Term $I_2$ (boundary).}\;
From~\eqref{eq:DFC-decomp}, $I_2 = \varphi_{j_2}\,\Pi_{j_2}
- \varphi_{j_1}\,\Pi_{j_1-1}$. Since $\Pi_{j_1-1}$ and $\Pi_{j_2}$ are
bounded by $\Pi_{\max} := \sup_{j \in [j_1-1,\,j_2]}|\Pi_j|$:
\begin{equation}\label{eq:I2-bound}
  |I_2|
  \;\le\;
  \Pi_{\max}\,\bigl(|\varphi_{j_1}| + |\varphi_{j_2}|\bigr).
\end{equation}
Note: $\Pi_{\max}$ is controlled by the energy input.  In the
statistically stationary cascade, $\Pi_j \approx \varepsilon$ uniformly,
so $\Pi_{\max} \le C\,\varepsilon$.

\textbf{Term $I_3$ (commutator remainder).}\;
By (DFC3)~\eqref{eq:DFC3-concrete} directly:
\begin{equation}\label{eq:I3-bound}
  |I_3|
  \;=\;
  \biggl|\sum_{j=j_1}^{j_2} \varphi_j\,R_j\biggr|
  \;\le\;
  \sum_{j=j_1}^{j_2} |\varphi_j\,R_j|
  \;\le\;
  C_{\mathrm{rem}}\,\bigl(|\varphi_{j_1}| + |\varphi_{j_2}|\bigr).
\end{equation}
No further estimates are needed: (DFC3) is stated as exactly the
inequality used here.

\medskip\noindent
\textbf{Combining.}\;
$\sum_{j=j_1}^{j_2} \varphi_j\,T_j
= I_1 + I_2 + I_3
\le \underbrace{-\Pi_0\,(\varphi_{j_1}-\varphi_{j_2})}_{\le\,0}
+ \underbrace{(\Pi_{\max} + C_{\mathrm{rem}})}_
  {=:\,C_{\mathrm{bdry}}}
\,(|\varphi_{j_1}|+|\varphi_{j_2}|)$.
This is exactly~\eqref{eq:NLw} with
$C_{\mathrm{bdry}} = \Pi_{\max} + C_{\mathrm{rem}}$.
\end{proof}

\begin{remark}[Edge cases and summation conventions]
\label{rem:edge-cases}
\mbox{}
\begin{itemize}
  \item \textbf{Window width $j_2 - j_1 \ge 2$:}\;
    The constraint ensures that the interior sum $I_1$ contains at least
    two terms ($j = j_1$ and $j = j_1+1$), so the telescoping
    $\sum_{j=j_1}^{j_2-1}(\varphi_j - \varphi_{j+1}) =
    \varphi_{j_1} - \varphi_{j_2}$ is non-trivial. For $j_2 - j_1 = 1$
    the sum $I_1$ has a single term and the bound degenerates to
    $|I_1| \le |\varphi_{j_1}-\varphi_{j_2}|\cdot\Pi_{\max}$, which is
    absorbed into $I_2$.
  \item \textbf{Summation bounds:}\;
    Interior sum: $j = j_1, \ldots, j_2-1$ (hence $j_2-j_1$ terms).
    Boundary term: $I_2 = \varphi_{j_2}\Pi_{j_2} -
    \varphi_{j_1}\Pi_{j_1-1}$ uses $\Pi_{j_1-1}$ (one index below the
    window). This is well-defined because the flux
    $\Pi_j$~\eqref{eq:flux} is defined for all~$j$.
  \item \textbf{Definition of $\Pi_{\max}$:}\;
    $\Pi_{\max} := \sup_{j \in [j_1-1,\, j_2]} |\Pi_j|$ (including the
    boundary index $j_1-1$). In the statistically stationary inertial
    range, $\Pi_j \approx \varepsilon$ uniformly, so $\Pi_{\max} \le
    (1+\delta)\,\varepsilon$ for some small $\delta$ depending on
    Reynolds number.
\end{itemize}
\end{remark}

\begin{proposition}[Spectral slope implies DFC2]\label{prop:slope-DFC2}
The sharp threshold for \textup{(DFC2)} is a dyadic ratio
$E_{j+1}/E_j \le 2^{-2/3}$ (i.e., a log-log slope $\le -2/3$), which
reflects the shell-width factor $\Delta k_j = k_j \ln 2$.  A convenient
\emph{sufficient} condition, directly matching the K41 prediction, is the
\emph{log-log slope condition}
\begin{equation}\label{eq:slope-condition}
  \frac{\mathrm{d}}{\mathrm{d}\ln k}\,\ln E(k)
  \;\le\; -\frac{5}{3}
  \qquad \text{for } k \in [k_{j_1},\, k_{j_2}],
\end{equation}
or equivalently in dyadic form: $E_{j+1}/E_j \le 2^{-5/3}$ for
$j \in [j_1, j_2-1]$.  Under this condition, \textup{(DFC2)} holds
with strict steepness $\varphi_j - \varphi_{j+1} \ge \ln 2 > 0$.
\end{proposition}

\begin{proof}
Since $E_j^* = C_K\,\varepsilon^{2/3}\,k_j^{-5/3}\,\Delta k_j$ with
$\Delta k_j = k_j \ln 2$ and $k_j = 2^j$, we have
$E_j^* = C_K\,\varepsilon^{2/3}\,(\ln 2)\,k_j^{-2/3}$, so the ratio
$E_j^*/E_{j+1}^* = (k_j/k_{j+1})^{-2/3} = 2^{2/3}$. The slope
condition $\frac{\mathrm{d}}{\mathrm{d}\ln k}\ln E(k) \le -5/3$
in dyadic form reads $E_{j+1}/E_j \le 2^{-5/3}$, which is
\emph{stronger} than the threshold $E_{j+1}/E_j \le 2^{-2/3}$
required for (DFC2). In particular,
$E_{j+1}/E_{j+1}^* = (E_{j+1}/E_j)\cdot(E_j^*/E_{j+1}^*)\cdot(E_j/E_j^*)
\le 2^{-5/3}\cdot 2^{2/3}\cdot(E_j/E_j^*) = 2^{-1}\cdot(E_j/E_j^*)
< E_j/E_j^*$, giving $\varphi_{j+1} < \varphi_j$.

\smallskip\noindent
\textbf{Sharp threshold.}\;
The minimal slope condition for (DFC2) is $E_{j+1}/E_j \le 2^{-2/3}$,
i.e., a log-log slope $\le -2/3$ (not $-5/3$). The stronger
condition~\eqref{eq:slope-condition} is stated because it directly
corresponds to the K41 prediction and is the form verified in DNS.
\end{proof}

\begin{remark}[Status of the Downhill Flux Condition]
\label{rem:DFC-status}
The DFC is a \emph{physically verifiable} condition, not a mathematical
assumption about the NS equations:
\begin{itemize}
  \item \textbf{(DFC1) Forward cascade:} Confirmed in all DNS at
    $R_\lambda \gtrsim 100$. Backscatter (localised inverse transfer)
    occurs but is sub-dominant: $\langle\Pi_j\rangle > 0$ throughout the
    inertial range~\cite{Frisch95,ES06,Kaneda03,Gotoh02}.
  \item \textbf{(DFC2) Spectral steepness:} The compensated spectrum
    $E(k)\,k^{5/3}/\varepsilon^{2/3}$ is approximately constant
    ($\approx C_K$) in the inertial range and \emph{decreasing} beyond
    the bottleneck. At $R_\lambda \gtrsim 200$, the ratio
    $E_j/E_j^*$ is non-increasing across the inertial range in all
    published DNS~\cite{ES06,Frisch95}. Deviations from monotonicity are
    confined to $\sim 2$ shells near the dissipation cutoff
    (bottleneck effect), which can be absorbed into the boundary
    constant $C_{\mathrm{bdry}}$.
  \item \textbf{(DFC3) Remainder control:} The
    inequality~\eqref{eq:DFC3-concrete} is a \emph{consequence}
    of two ingredients: (a)~the explicit commutator
    bound~\eqref{eq:remainder-bound} from~\cite{CET94,DR00}, and
    (b)~DFC2 monotonicity ($\sup_{[j_1,j_2]}|\varphi_j| \le
    |\varphi_{j_1}|$). Both are rigorous; no empirical input is
    needed. Concretely: $C_{\mathrm{rem}} =
    C_{\mathrm{rem}}^{\mathrm{univ}}\,\|u\|_{L^3_t
    B^{1/3}_{3,\infty}}^3$, and the Besov norm is bounded by
    $C\,\|u_0\|_{L^2}$ via the energy inequality and Bernstein's
    lemma~\cite{CET94}.
\end{itemize}

\medskip\noindent
\textbf{Dependency table.}
\medskip

\begin{center}
\begin{tabular}{llll}
\hline
\textbf{Label} & \textbf{Statement} & \textbf{Type} & \textbf{Source} \\
\hline
DFC1 & Forward cascade & Empirical (DNS) & \cite{Frisch95} \\
DFC2 & Spectral steepness & Empirical (DNS) & \cite{ES06} \\
DFC3 & Besov regularity & Lemma (conditional on $B^{1/3}_{3,\infty}$) & \cite{CET94} \\
Lem.~\ref{lem:DFC-implies-NLw} & DFC $\Rightarrow$ NL$'$ & \textbf{Theorem} & This paper \\
Prop.~\ref{prop:window-NL} & NL$'$ $\Rightarrow$ Thm~\ref{thm:anomalous} & \textbf{Theorem} & This paper \\
\hline
\end{tabular}
\end{center}

\medskip\noindent
The logical chain is:
$\textup{DFC1} + \textup{DFC2} + \textup{DFC3}
\;\Longrightarrow{\text{Lem.~\ref{lem:DFC-implies-NLw}}}\;
\textup{NL}'
\;\Longrightarrow{\text{Prop.~\ref{prop:window-NL}}}\;
\textup{Thm~\ref{thm:anomalous}}$.
The only non-rigorous inputs are (DFC1) and (DFC2), stated as
\emph{falsifiable inequalities} on the measured spectral flux
($\Pi_j \ge \Pi_0$) and compensated slope ($\varphi_{j+1} \le
\varphi_j$). The implication $\textup{DFC} \Rightarrow \textup{NL}'
\Rightarrow \textup{Thm~\ref{thm:anomalous}}$ is purely analytic.
A reviewer can verify or falsify (DFC1)--(DFC2) from published DNS
data independently of the mathematical framework.
\end{remark}

\begin{remark}[Independent support for DFC1 from vanishing viscosity theory]
\label{rem:Drivas-VV}
The forward-cascade assumption (DFC1) receives independent theoretical
support from the vanishing viscosity programme of
Drivas~\cite{Drivas2019cascade}:
for any space-time $L^3$ weak solution of the Euler equations arising as a
strong limit of $d\ge 3$-dimensional Navier--Stokes solutions, the
Duchon--Robert defect measure $D(u)$ is non-negative, and the Lagrangian
forward/backward dispersion measure matches the energy defect in the sense
of distributions. The resulting Lagrangian time-asymmetry (particles
disperse faster backward than forward in $d\ge 3$) is a rigorous
characterisation of the forward cascade direction.

De~Rosa, Drivas and Inversi~\cite{DeRosaDrivasInversi2024} sharpen this
further: any bounded Euler solution arising as a zero-viscosity limit of
Navier--Stokes solutions cannot have anomalous dissipation supported on a
set of Hausdorff dimension smaller than~$d$. This bound is sharp.

These results do \emph{not} rigorise DFC1 in its present form
($\Pi_j \ge \Pi_0 > 0$, a strict lower bound on the shell-averaged flux),
but they provide a mathematically rigorous justification for the physical
expectation $D(u) \ge 0$ that underlies it. The strict positivity of
$\Pi_j$ across the inertial range remains an empirical input.
\end{remark}

The following two lemmas provide the analytic backbone. The first
isolates the commutator remainder bound as a self-contained estimate;
the second gives the flux decomposition that drives the entire argument.

\begin{lemma}[Commutator remainder bound]\label{lem:remainder}
Let $u$ be a Leray--Hopf weak solution on $\mathbb{T}^3$ and let $R_j$
be the Littlewood--Paley commutator remainder from the decomposition
$T_j = \Pi_{j-1} - \Pi_j + R_j$.
\begin{enumerate}
  \item[\textup{(i)}] $\sum_j R_j = 0$ (conservation).
  \item[\textup{(ii)}] If $u \in L^3_t B^{1/3}_{3,\infty}(\mathbb{T}^3)$
    with norm $\le B$, then for any sub-window $[j_1, j_2]$ and any
    bounded weights $(\varphi_j)_j$:
    \begin{equation}\label{eq:Rj-bound}
      \sum_{j=j_1}^{j_2} |\varphi_j|\,|R_j|
      \;\le\;
      C_0\,B^3\,\sup_{j \in [j_1,j_2]} |\varphi_j|,
    \end{equation}
    where $C_0 > 0$ is a universal constant.
  \item[\textup{(iii)}] Every Leray--Hopf solution with
    $\|u_0\|_{L^2} \le M$ satisfies $B \le C_1\,M$ (energy inequality
    $+$ Bernstein's lemma).
\end{enumerate}
\end{lemma}

\begin{proof}
Part~(i) follows from $\sum_j T_j = 0$ and $\sum_j(\Pi_{j-1}-\Pi_j)=0$
(telescoping). Part~(ii) is the standard paraproduct/commutator estimate:
each $R_j$ collects terms of the form $[\Delta_j, S_{j'}](u \otimes u)$,
and the off-diagonal decay $\|[\Delta_j, S_{j'}]\|_{L^1 \to L^1}
\le C\,2^{-|j-j'|/3}$ combined with summing over $j'$ gives $\|R_j\|_{L^1}
\le C_0\,B^3$; see~\cite{CET94}, Proposition~1 and~\cite{DR00},
Lemma~2.1. The weighted sum follows from
$\sum |{\varphi_j}|\,|{R_j}| \le \sup|\varphi_j|\,\sum|R_j|
\le \sup|\varphi_j|\cdot C_0 B^3$.
Part~(iii): the energy inequality $\frac{1}{2}\|u(t)\|_{L^2}^2
+ \nu\int_0^t \|\nabla u\|_{L^2}^2 \le \frac{1}{2}\|u_0\|_{L^2}^2
+ \int_0^t \langle f,u\rangle$ combined with Bernstein's inequality
$\|\Delta_j u\|_{L^3} \le C\,2^{j/3}\|\Delta_j u\|_{L^2}$ gives the
Besov bound.
\end{proof}

The following lemma, based on the distributional energy balance of
Duchon and Robert~\cite{DR00}, provides the flux decomposition.

\begin{lemma}[Flux representation]\label{lem:flux}
For any family of weights $(\varphi_j)_j$ with $\sup_j |\varphi_j| < \infty$,
the nonlinear transfer admits the decomposition
\begin{equation}\label{eq:flux_rep}
  \sum_j \varphi_j\,T_j
  \;=\;
  \sum_j (\varphi_{j+1} - \varphi_j)\,\Pi_j
  \;+\;
  \sum_j \varphi_j\,R_j,
\end{equation}
where $\Pi_j$ is the energy flux through shell~$j$
(eq.~\eqref{eq:flux}) and the commutator remainder $R_j$ collects the
Littlewood--Paley paraproduct commutator terms; $\sum_j R_j = 0$.
For any sub-window $[j_1, j_2]$, the weighted remainder satisfies the
\emph{explicit bound}
\begin{equation}\label{eq:remainder-bound}
  \sum_{j=j_1}^{j_2} |\varphi_j\,R_j|
  \;\le\;
  C_{\mathrm{rem}}\,\|u\|_{L^3_t B^{1/3}_{3,\infty}}^3
  \;\cdot\;
  \sup_{j \in [j_1,j_2]} |\varphi_j|,
\end{equation}
where $C_{\mathrm{rem}} > 0$ is a universal constant (independent of the
window, the solution, and the reference spectrum).  This follows from the
standard paraproduct decomposition $R_j = \sum_{\ell} [T_\ell,
\Delta_j]$ and the commutator estimate $\|[T_\ell, \Delta_j]\|_{L^1}
\le C\,2^{-|j-\ell|/3}\,\|u\|_{B^{1/3}_{3,\infty}}^3$
(see~\cite{CET94}, Proposition~1; \cite{DR00}, Lemma~2.1).
\end{lemma}

\begin{proof}[Proof (sketch)]
Write $T_j = \Pi_{j-1} - \Pi_j + R_j$, where
$\Pi_j = -\langle \Delta_{\le j} u \otimes \Delta_{\le j} u,\,
\nabla \Delta_{\le j} u \rangle$ is the cumulative flux
and $R_j$ collects the Littlewood--Paley commutator terms.
Abel summation gives~\eqref{eq:flux_rep}.  The commutator estimate
follows from the Duchon--Robert distributional energy balance and
standard paraproduct estimates in Besov spaces; see~\cite{DR00}
and~\cite{CET94}.
\end{proof}

\begin{theorem}[Anomalous dissipation bound]\label{thm:anomalous}
Let $(u^\nu)_{\nu > 0}$ be a family of Leray--Hopf weak solutions of the
incompressible Navier--Stokes equations on $\mathbb{T}^3$:
\begin{equation}\label{eq:NSE}
  \partial_t u^\nu + (u^\nu \cdot \nabla)\,u^\nu
  + \nabla p^\nu
  = \nu\,\Delta u^\nu + f,
  \qquad \nabla \cdot u^\nu = 0,
\end{equation}
with a fixed smooth forcing $f$ and initial data satisfying
$\|u^\nu_0\|_{L^2} \le M$.  Denote the time-averaged dissipation rate
\begin{equation}\label{eq:diss}
  \varepsilon_\nu
  \;:=\;
  \frac{\nu}{T} \int_0^T \!\!\int_{\mathbb{T}^3}
    |\nabla u^\nu|^2 \,dx\,dt.
\end{equation}
Assume that the cascade satisfies \emph{one} of the following
(in decreasing order of strength):
\begin{enumerate}
  \item[\textup{(H1)}] Entropy compatibility~\eqref{eq:NL}
    (Definition~\ref{def:NL}), or
  \item[\textup{(H2)}] Window-entropy compatibility~\eqref{eq:NLw}
    (Definition~\ref{def:NLw}), or
  \item[\textup{(H3)}] The Downhill Flux Condition
    \textup{(DFC1)--(DFC3)} of Definition~\ref{def:DFC}.
\end{enumerate}
Further assume \textup{($\nu$-uniform energy injection)}:
there exists $m_0 > 0$, independent of $\nu$, such that
$T^{-1}\int_0^T \langle f, u^\nu \rangle\,dt \ge m_0$
for all $\nu > 0$.

Then there exists a constant $\varepsilon_* > 0$, depending only on $f$
and $M$, such that
\begin{equation}\label{eq:anomalous}
  \liminf_{\nu \to 0}\; \varepsilon_\nu
  \;\ge\; \varepsilon_*.
\end{equation}
\end{theorem}

\begin{proof}[Proof (sketch)]
\textbf{Step 1: Energy balance.}\;
Multiplying~\eqref{eq:NSE} by $u^\nu$ and integrating yields the global
energy balance
\begin{equation}\label{eq:energy_balance}
  \frac{d}{dt}\frac{1}{2}\|u^\nu\|_{L^2}^2
  \;=\;
  -\nu\|\nabla u^\nu\|_{L^2}^2
  \;+\;
  \langle f, u^\nu \rangle.
\end{equation}
In the statistically stationary state, the time-averaged dissipation
equals the time-averaged energy input:
$\varepsilon_\nu = T^{-1}\int_0^T \langle f, u^\nu \rangle\,dt$.

\textbf{Step 2: Regularised free-energy monotonicity.}\;
For $\delta > 0$, define the regularised free energy
\[
  \mathcal{F}_\delta[\mathbf{E}]
  \;:=\;
  \sum_j \Bigl[
    (E_j + \delta)\,\ln\frac{E_j + \delta}{E_j^* + \delta}
    \;-\; (E_j - E_j^*)
  \Bigr].
\]
This is $C^1$ in all arguments and satisfies $\mathcal{F}_\delta \ge 0$
with equality iff $\mathbf{E} = \mathbf{E}^*$.
Computing $\frac{d}{dt}\mathcal{F}_\delta$ along the shell-energy
evolution $\frac{1}{2}\frac{d}{dt}E_j = -\nu k_j^2 E_j + T_j
+ \langle \Delta_j u, \Delta_j f\rangle$ (up to commutator corrections),
the three contributions are:
\begin{enumerate}
  \item \emph{Viscous term:} contributes
    $-2\nu \sum_j k_j^2\,E_j^\nu\,
    \ln\!\bigl(\frac{E_j^\nu+\delta}{E_j^*+\delta}\bigr)$.
    Individual summands $k_j^2\,E_j\,\ln(E_j/E_j^*)$ can have either
    sign (positive when $E_j > E_j^*$, negative when $E_j < E_j^*$),
    so the viscous contribution does not have a definite sign.
    Its role in the monotonicity~\eqref{eq:F_decay} is to provide a
    dissipative correction that drives the spectrum toward~$\mathbf{E}^*$:
    shells with $E_j > E_j^*$ are dissipated faster than equilibrium,
    while shells with $E_j < E_j^*$ are dissipated more slowly;
  \item \emph{Nonlinear transfer:} by Assumption~\eqref{eq:NL}
    (entropy-compatible cascade),
    $\sum_j \ln\bigl(\frac{E_j+\delta}{E_j^*+\delta}\bigr)\,T_j
    \le 0$;
  \item \emph{Forcing:} bounded by $C\,\|f\|_{H^s}^2$ via Cauchy--Schwarz.
\end{enumerate}
Combining these and passing $\delta \downarrow 0$ via Fatou's lemma
(lower semicontinuity of $\mathcal{F}$) yields
\begin{equation}\label{eq:F_decay}
  \frac{d}{dt}\,\mathcal{F}[\mathbf{E}^\nu(t)]
  \;\le\;
  -\,2\nu \sum_{j} k_j^2\,
    E_j^\nu\,\ln\!\frac{E_j^\nu}{E_j^*}
  \;+\;
  C\,\|f\|_{H^s}^2.
\end{equation}
The quantity $D_\mathcal{F}^\nu := \sum_j k_j^2\,
E_j^\nu\,\ln(E_j^\nu/E_j^*)$ is the \emph{entropy dissipation}.
It is not sign-definite (individual terms can be positive or negative).
Each summand $k_j^2\,E_j\,\ln(E_j/E_j^*)$ vanishes if and only if
$E_j = E_j^*$ (since $\ln(x/y) = 0$ iff $x = y$ for $x,y > 0$,
and $k_j^2\,E_j > 0$); hence $D_\mathcal{F}^\nu = 0$ with all
summands individually zero if and only if $\mathbf{E}^\nu = \mathbf{E}^*$.
The dissipative character of the viscous term is encoded in the
full inequality~\eqref{eq:F_decay}: combined with the non-positive
nonlinear contribution and the bounded forcing, it ensures that
$\mathcal{F}$ decreases on average.

\textbf{Step 3: Lower bound on dissipation (using ND).}\;
Under the $\nu$-uniform energy injection assumption (stated in Step~4
below), the energy balance (Step~1) yields $\varepsilon_\nu \ge m_0 > 0$
directly. Without this assumption, one can only argue heuristically:
if $\varepsilon_\nu \to 0$, then for smooth, non-trivial $f$ and bounded
$\|u^\nu\|_{L^2} \le C(M)$, the forcing term $\langle f, u^\nu \rangle$
remains bounded but the dissipation vanishes, which is inconsistent
with sustained energy injection. The ND assumption formalises this.

\textbf{Step 4: Quantitative bound.}\;
The $\mathcal{F}$-monotonicity~\eqref{eq:F_decay} controls the
entropy dissipation $D_\mathcal{F}^\nu(t) := \sum_j k_j^2\,
E_j^\nu\,\ln(E_j^\nu/E_j^*)$, which measures the spectral distance
from the Kolmogorov reference~$\mathbf{E}^*$.
However, the physical dissipation $\varepsilon_\nu = \nu\sum_j k_j^2 E_j^\nu$
is not directly controlled by~$D_\mathcal{F}^\nu$ without an additional
\emph{non-degeneracy} assumption linking these two quantities.

\emph{Assumption ND ($\nu$-uniform energy injection).}\;
There exists $m_0 > 0$, independent of $\nu$, such that
\begin{equation}\label{eq:ND}
  \frac{1}{T}\int_0^T \langle f, u^\nu \rangle\,dt \;\ge\; m_0 > 0
  \qquad \text{for all } \nu > 0.
\end{equation}
Under this assumption, the energy balance (Step~1) immediately gives
$\varepsilon_\nu \ge m_0$ for all~$\nu$, hence
$\liminf_{\nu\to 0}\,\varepsilon_\nu \ge m_0 > 0$.

A refined bound can be obtained as follows.  For divergence-free
$f \in H^{-1}(\mathbb{T}^3)$ with $\langle f, \cdot\rangle$ non-trivial
on divergence-free fields, the Poincar\'e inequality on $\mathbb{T}^3$
gives $\|u^\nu\|_{L^2} \le C_P\,\|\nabla u^\nu\|_{L^2}$, so that
$\langle f, u^\nu \rangle
\le \|f\|_{H^{-1}}\,\|\nabla u^\nu\|_{L^2}
\le \|f\|_{H^{-1}}\,(\varepsilon_\nu/\nu)^{1/2}$.
Combining with~\eqref{eq:ND}:
$m_0 \le \|f\|_{H^{-1}}\,(\varepsilon_\nu/\nu)^{1/2}$, which gives
$\varepsilon_\nu \ge \nu\,m_0^2/\|f\|_{H^{-1}}^2$.  This bound is
non-trivial only for fixed~$\nu$ and degenerates as $\nu \to 0$.
The $\nu$-\emph{independent} lower bound
$\varepsilon_* \ge m_0$ from Step~1 remains the operative estimate;
a sharper $\nu$-independent bound of the form
$\varepsilon_* \ge c\,\|f\|_{H^{-1}}^2 / M^2$
would require a spectral Poincar\'e inequality
linking $D_\mathcal{F}^\nu$ to~$\varepsilon_\nu$, which is an open
problem (see the Remark below).
The natural Sobolev norm for the energy injection is~$H^{-1}$ (dual
to~$\|\nabla u\|_{L^2}$); the $H^s$ with $s > -1$ appearing
in~\eqref{eq:F_decay} arises from the shell-wise forcing decomposition
and is strictly weaker.

% (Full proof details in the expanded version.)
\end{proof}

\begin{remark}[Role of the free-energy machinery]
The quantitative bound $\varepsilon_* \ge m_0$ in Step~4 follows
directly from the energy balance and assumption~ND, without invoking
the functional~$\mathcal{F}$.  The non-trivial contribution of the
free-energy approach lies in \emph{Steps~1--2}: the
$\mathcal{F}$-monotonicity~\eqref{eq:F_decay} provides a
\emph{structural} explanation for why the K41 profile is the attracting
state of the cascade, and it controls the entropy dissipation
$D_\mathcal{F}^\nu$, which measures spectral distance from K41.
The quantitative anomalous-dissipation bound can be sharpened (beyond
the crude $\varepsilon_\nu \ge m_0$) once the entropy-dissipation
$D_\mathcal{F}^\nu$ is related to the physical dissipation via a
\emph{spectral Poincar\'e inequality}.  Establishing such an inequality
for Leray--Hopf solutions is an open problem.
\end{remark}

\begin{remark}[Scope and validity of the entropy-compatibility hypothesis]
Theorem~\ref{thm:anomalous} is conditional on
Assumption~\eqref{eq:NL}.  Three settings in which this hypothesis
is known or expected to hold:
\begin{enumerate}
\item \emph{Shell models:} In GOY and Sabra shell models with a
  Markov-type transfer structure, the nonlinear term is a genuine
  mass-preserving operator on shell energies, and~\eqref{eq:NL}
  follows from the standard entropy-dissipation inequality for
  relative entropies under doubly stochastic maps.
\item \emph{Stochastic cascade models:} In multiplicative cascade models
  of turbulence (Mandelbrot, Kahane), the energy redistribution is
  by construction a Markov kernel, yielding~\eqref{eq:NL}.
\item \emph{Navier--Stokes with downhill flux:} For NS solutions whose
  energy flux $\Pi_j$ satisfies a ``downhill'' condition---energy flows
  preferentially from overexcited to underexcited shells relative
  to~$\mathbf{E}^*$---Lemma~\ref{lem:flux} combined with Abel
  summation gives~\eqref{eq:NL} up to controllable commutator
  remainders.  Whether this downhill property holds generically for
  NS solutions in the inertial range is an open question closely
  related to the Onsager conjecture and the structure of the
  Duchon--Robert defect measure~\cite{DR00}.
\end{enumerate}
\end{remark}

% ===================================================================
\section{Intermittency Corrections}\label{sec:intermittency}

The K41 theory predicts that the $p$-th order longitudinal structure
function scales as
\begin{equation}\label{eq:Sp_K41}
  S_p(\ell) \;:=\;
  \bigl\langle |\delta_\ell u|^p \bigr\rangle
  \;\sim\; (\varepsilon\,\ell)^{p/3},
  \qquad
  \zeta_p^{\mathrm{K41}} = \frac{p}{3}.
\end{equation}
Experiments reveal systematic deviations: $\zeta_p < p/3$ for $p > 3$,
a signature of \emph{intermittency}.

\subsection{Fluctuations around the variational minimum}

In our framework, intermittency arises from fluctuations of the shell
energies $\mathbf{E}$ around the K41 minimiser $\mathbf{E}^*$.  Expanding
$\mathcal{F}$ to second order:
\begin{equation}\label{eq:F_expand}
  \mathcal{F}[\mathbf{E}^* + \delta\mathbf{E}]
  \;=\;
  \mathcal{F}[\mathbf{E}^*]
  \;+\;
  \frac{1}{2}\sum_{j}
    \frac{(\delta E_j)^2}{E_j^*}
  \;+\; O(|\delta\mathbf{E}|^3).
\end{equation}
The Hessian $H_{jj} = 1/E_j^*$ implies that fluctuations at small scales
(large $j$, small $E_j^*$) are suppressed more strongly than fluctuations
at large scales.  This asymmetry is the variational origin of
intermittency.

\subsection{Connection to the K62 log-normal model}

If we model the fluctuations $\delta E_j$ as Gaussian in the variable
$\ln(E_j/E_j^*)$, the resulting structure function exponents are
\begin{equation}\label{eq:zeta_p}
  \zeta_p
  \;=\;
  \frac{p}{3}
  \;-\;
  \frac{\mu}{18}\,p(p-3),
\end{equation}
which is exactly the K62 (Kolmogorov--Obukhov) refined similarity
hypothesis with intermittency parameter $\mu$~\cite{K62,Frisch95}.  The
parameter $\mu$ is determined by the variance of $\ln(E_j/E_j^*)$ under
the Gibbs measure $\exp(-\mathcal{F}/\Theta)$, where the effective
temperature $\Theta > 0$ parametrises the strength of turbulent
fluctuations around the K41 equilibrium (analogous to the temperature
in a canonical ensemble).  In the limit $\Theta \to 0$, the Gibbs
measure concentrates on the K41 minimiser and $\mu \to 0$ (no
intermittency); for finite $\Theta$, the log-normal variance scales as
$\mu \propto \Theta\,E_j^*$, connecting the intermittency parameter
to the intensity of the cascade fluctuations.

\begin{remark}
The She--L\'ev\^eque model~\cite{SL94} proposes the alternative
\[
  \zeta_p = \frac{p}{9} + 2\Bigl(1 - \bigl(\tfrac{2}{3}\bigr)^{p/3}\Bigr),
\]
which fits experimental data more accurately for high $p$.  Deriving this
from the free-energy framework would require modelling the fluctuations as
a log-Poisson rather than log-normal process.  This is an open problem
within our approach.
\end{remark}

% ===================================================================
\subsection{Structural Classification}\label{sec:resolvent-cascade}

The joint minimisation of Theorem~\ref{thm:K41} exhibits a three-level
hierarchy: (i)~the leading-order energy budget is neutral (any power-law
spectrum is kinematically admissible); (ii)~the first-order constant-flux
constraint restricts but does not uniquely determine the profile;
(iii)~the second-order strict convexity of~$\mathcal{F}$ selects K41 as
the unique minimiser.  This pattern---referred to as
\emph{second-order resolvent dominance} in~\cite{FST-Unified2026}---has
analogues in other variational selection problems and is discussed in
detail there.

% ===================================================================
\subsection{Robustness and game-theoretic structure}\label{sec:robustness}

The strict-DFC framework of Definition~\ref{def:DFC} requires
pointwise lower bounds $\Pi_j \ge \Pi_0$ and exact monotonicity of the
compensated spectrum.  In practice, DNS data exhibit small-scale
fluctuations around these idealisations.  We formulate two extensions
that make the variational principle robust and reveal its game-theoretic
content.

\begin{proposition}[Relaxed flux constraint]
\label{prop:dfc-slack}
Let $\{\eta_j\}_{j=1}^{N-1}$ and $\{\delta_j\}_{j=1}^{N-1}$ be
non-negative sequences satisfying $\sum_j \eta_j < \infty$ and
$\sum_j \delta_j < \infty$.  Define the \emph{relaxed admissible set}
\[
  \mathcal{A}_{\textup{slack}}
  \;=\;
  \bigl\{
    (\Pi_j)_{j=1}^{N-1} :\;
    \Pi_j \ge \Pi_0 - \eta_j,\quad
    \varphi_{j+1} \le \varphi_j + \delta_j
    \;\text{for all } j
  \bigr\},
\]
where $\varphi_j = F_j[\mathbf{E}] - F_j[\mathbf{E}^*]$ denotes the
shell-wise KL divergence.  Then the minimiser of
$\mathcal{F}[\mathbf{E}]$ over $\mathcal{A}_{\textup{slack}}$
satisfies
\begin{equation}\label{eq:dfc-slack-bound}
  \|\mathbf{E} - \mathbf{E}^*\|_{\ell^2}^2
  \;\le\;
  C\,\Bigl(\sum_j \eta_j \;+\; \sum_j \delta_j\Bigr)
\end{equation}
for a constant $C$ depending only on the spectral gap of the
strict-DFC problem.  In particular, K41 is recovered as
$\eta_j, \delta_j \to 0$.
\end{proposition}

\begin{proof}[Proof sketch]
The strict-DFC functional $\mathcal{F}$ restricted to the exact
admissible set $\mathcal{A}$ has a unique minimiser $\mathbf{E}^*$
with positive spectral gap $\gamma > 0$ (i.e.\
$\mathcal{F}[\mathbf{E}] \ge \gamma\,\|\mathbf{E} -
\mathbf{E}^*\|_{\ell^2}^2$ near $\mathbf{E}^*$).  The relaxed set
$\mathcal{A}_{\textup{slack}} \supset \mathcal{A}$ enlarges the
feasible region by perturbations of total size
$\sigma := \sum_j \eta_j + \sum_j \delta_j$.  By standard
perturbation theory for convex programmes (cf.\
Bonnans--Shapiro~\cite{BonnansShapiro00}), the minimiser over
$\mathcal{A}_{\textup{slack}}$ shifts by at most
$O(\sigma/\gamma)$ in~$\ell^2$, and the objective value changes by
at most $O(\sigma^2/\gamma)$.  The bound~\eqref{eq:dfc-slack-bound}
follows with $C = 1/\gamma$.
\end{proof}

\begin{remark}
Proposition~\ref{prop:dfc-slack} shows that the K41 selection is
\emph{stable}: arbitrarily small summable perturbations of the flux
constraint and the monotonicity condition produce correspondingly
small deviations from the Kolmogorov spectrum.  This is essential for
the physical relevance of the framework, since no finite-Reynolds-number
flow satisfies the DFC exactly.
\end{remark}

\begin{proposition}[Minimax characterisation]
\label{prop:minimax}
The K41 spectrum $\mathbf{E}^*$ is the saddle point of the minimax
problem
\begin{equation}\label{eq:minimax}
  \mathbf{E}^*
  \;=\;
  \arg\min_{\mathbf{E} \in \mathcal{E}}\;
  \max_{\boldsymbol{\Pi} \in \mathcal{A}}\;
  \Bigl[
    \mathcal{F}[\mathbf{E}]
    \;-\;
    \lambda \sum_j (\Pi_j - \varepsilon)^2
  \Bigr],
\end{equation}
where $\mathcal{E}$ is the set of non-negative energy spectra with
fixed total energy, and $\mathcal{A}$ is the set of admissible flux
profiles.  The min-player (viscous dissipation) selects the spectrum,
while the max-player (nonlinear transfer) selects the flux profile.
\end{proposition}

\begin{proof}[Proof sketch]
The objective in~\eqref{eq:minimax} is convex in~$\mathbf{E}$ (since
$\mathcal{F}$ is strictly convex and the penalty term is independent
of~$\mathbf{E}$) and concave in~$\boldsymbol{\Pi}$ (since the penalty
$-\lambda\sum_j(\Pi_j - \varepsilon)^2$ is concave).  Both
$\mathcal{E}$ and $\mathcal{A}$ are convex and compact (in the
finite-dimensional truncation to $N$ shells).  By Sion's minimax
theorem, the minimax value equals the maximin value, and a saddle point
exists.  The Karush--Kuhn--Tucker conditions at the saddle point
reproduce:
\begin{enumerate}[label=(\roman*)]
  \item the Euler--Lagrange equation
    $\partial\mathcal{F}/\partial E_j = \mu$ (constant Lagrange
    multiplier), which yields $E_j = E_j^*$ (K41);
  \item the stationarity in~$\Pi_j$:
    $\Pi_j = \varepsilon$ for all~$j$ (constant flux).
\end{enumerate}
The uniqueness of the saddle point follows from the strict
convexity--concavity of the objective.
\end{proof}

\begin{remark}
The minimax formulation reveals the \emph{game-theoretic structure}
of the turbulent cascade.  Viscous dissipation acts as a conservative
agent minimising the free energy (favouring smooth, large-scale
motion), while the nonlinear cascade acts as an adversarial agent
maximising the energy transfer toward small scales.  The K41 spectrum
is the Nash equilibrium of this two-player zero-sum game.
This perspective connects the variational framework to the theory of
adversarial robustness in optimisation~\cite{Rockafellar70} and
suggests that deviations from K41 can be understood as exploitable
strategies by one player when the other is constrained.
\end{remark}

\begin{remark}[Wasserstein convexity and Talagrand bound]
\label{rem:talagrand}
The functional $F[E]$ is uniformly convex in the $\ell^2$-Wasserstein metric on the space of energy spectra, with convexity constant $\lambda_F$ determined by the KL contribution. By the Otto--Villani theorem, the K41 minimiser $E^*$ attracts all admissible spectra at rate $W_2(E, E^*) \leq \sqrt{2 F[E] / \lambda_F}$. This provides a stability guarantee independent of Reynolds number, complementing the DFC-slack bound (Proposition~\ref{prop:dfc-slack}).
\end{remark}

% ===================================================================
\subsection{Open conjectures and intrinsic characterisation}\label{sec:open-conjectures}

The variational framework developed above suggests several precise conjectures
that connect the free-energy structure to classical turbulence phenomenology.

\begin{conjecture}[Spectral Poincar\'e inequality]
\label{conj:spectral-poincare}
For the energy spectrum $E(k)$ of a statistically stationary turbulent flow at viscosity $\nu$, the entropy dissipation $D_F^\nu = -dF[E]/dt$ and the physical dissipation $\varepsilon_\nu = 2\nu \int k^2 E(k)\, dk$ are related by a spectral Poincar\'e inequality:
\[
D_F^\nu \geq \kappa_\nu \cdot \varepsilon_\nu,
\]
where $\kappa_\nu > 0$ is the spectral Poincar\'e constant. If $\kappa_\nu$ remains bounded below as $\nu \to 0$ (i.e., $\kappa_\nu \geq \kappa_0 > 0$), then the K41 minimiser is dynamically stable in the inviscid limit: perturbations of $E^*$ are dissipated at a rate proportional to the physical dissipation.
\end{conjecture}

\begin{remark}
The spectral Poincar\'e inequality bridges the thermodynamic (entropy-based) and physical (energy-based) descriptions of turbulence. Its validity would imply that the K41 spectrum is not merely a static minimiser but an attractor of the turbulent dynamics.
\end{remark}

\subsubsection*{Threshold axiom and diagnostic framework}

The spectral Poincar\'e inequality admits a precise formulation as a \emph{threshold axiom}---the minimal structural condition that separates dissipative cascade dynamics from anomalous trapping.

\begin{axiom}[Spectral Poincar\'e -- SP-TU]\label{ax:sp-tu}
Let $\mathcal{H} = \{E = (E_j)_{j \geq 1} : E_j \geq 0,\, \sum E_j < \infty\}$ be the space of shell energy distributions, let $\mathcal{N} \subset \mathcal{H}$ be the subspace of null modes (conservation laws: total energy $\sum E_j = \mathrm{const}$), and let $\mu_{\mathcal{F}}$ be the Gibbs measure $\mu_{\mathcal{F}} \propto e^{-\mathcal{F}[E]/\Theta}$. There exists a spectral gap $\lambda > 0$ such that
\begin{equation}\label{eq:sp-tu}
  \mathrm{Var}_{\mu_{\mathcal{F}}}(g) \leq \frac{1}{\lambda}\, \mathcal{E}_{\mathcal{F}}(g, g) \qquad \text{for all } g \perp \mathcal{N},
\end{equation}
where $\mathcal{E}_{\mathcal{F}}(g,g) = \int |\nabla g|^2\, d\mu_{\mathcal{F}}$ is the Dirichlet form associated with the free-energy landscape.
\end{axiom}

\begin{proposition}[Equivalences of SP-TU]\label{prop:sp-equivalences}
The following are equivalent:
\begin{enumerate}
\item[(i)] \textbf{Spectral gap:} Axiom~\ref{ax:sp-tu} holds with $\lambda > 0$.
\item[(ii)] \textbf{Exponential relaxation:} For the semigroup $P_t$ of the energy cascade dynamics,
  $\mathrm{Var}_{\mu_{\mathcal{F}}}(P_t g) \leq e^{-2\lambda t}\, \mathrm{Var}_{\mu_{\mathcal{F}}}(g)$ for all $g \perp \mathcal{N}$.
\item[(iii)] \textbf{Rayleigh characterisation:}
  $\lambda = \inf_{g \perp \mathcal{N},\, g \neq 0} \frac{\mathcal{E}_{\mathcal{F}}(g,g)}{\mathrm{Var}_{\mu_{\mathcal{F}}}(g)}$.
\end{enumerate}
\end{proposition}

\begin{proof}[Proof sketch]
(i)$\Leftrightarrow$(iii) is the standard variational characterisation of the spectral gap. (i)$\Rightarrow$(ii) follows from the spectral theorem for self-adjoint generators; (ii)$\Rightarrow$(i) by differentiating at $t=0$.
\end{proof}

\begin{remark}[Diagnostic interface]\label{rem:sp-diagnostics}
The Rayleigh characterisation~(iii) provides both analytical and numerical access to $\lambda$:
\begin{itemize}
\item \textbf{Analytical:} Test functions of the form $g_j = \ln(E_j/E_j^*)$ yield $\mathcal{E}_{\mathcal{F}}(g_j, g_j) / \mathrm{Var}(g_j) = 1/E_j^*$, so $\lambda \geq \min_j (1/E_j^*) = 1/(C_K \varepsilon^{2/3} k_1^{-2/3} \Delta k_1) > 0$. This is the diagonal Hessian bound (Remark~\ref{rem:hessian}).
\item \textbf{Numerical:} Estimate $\lambda$ from the exponential decay of autocorrelations $\langle g(t)\, g(0) \rangle_{\mu_{\mathcal{F}}} \sim e^{-\lambda t}$ in shell-model simulations (cf.\ Remark~\ref{rem:sabra-dfc}).
\item \textbf{Block decomposition:} Establish local Poincar\'e on shell clusters $[j_1, j_2]$ and glue via controlled overlap (Zegarlinski block dynamics, analogous to the Yang--Mills strategy in the companion paper \cite{Geiger2026-YM}).
\end{itemize}
\end{remark}

\begin{remark}[No-go mechanisms]\label{rem:sp-nogo}
SP-TU fails if and only if there exist \emph{quasi-null modes}: a sequence $g_k \perp \mathcal{N}$ with $\mathrm{Var}(g_k) = 1$ and $\mathcal{E}(g_k, g_k) \to 0$. Three physical mechanisms can produce such modes:
\begin{enumerate}
\item \textbf{Metastability:} The cascade dynamics traps in nearly invariant regions (e.g., bottleneck accumulation), creating slow indicator observables with vanishing Dirichlet form.
\item \textbf{Multi-scale leakage:} Dissipation acts on small scales while variance ``parks'' on large scales, producing a defective coercivity that cannot be closed by a single $\lambda$.
\item \textbf{Intermittency bursts:} Rare, intense events create heavy-tailed fluctuations that violate the Gaussian assumption underlying the Hessian bound.
\end{enumerate}
For K41 turbulence in the inertial range, none of these mechanisms is operative: the forward cascade (DFC1) prevents trapping, the scale locality (A1) prevents leakage, and the log-normal intermittency (Section~\ref{sec:intermittency}) decays as $1/N_{\mathrm{shells}}$ (Remark~\ref{rem:mu-disambiguation}). SP-TU is therefore \emph{expected} to hold, but a rigorous proof remains open.
\end{remark}

\begin{lemma}[SP-TU implies universal dissipation control]\label{lem:sp-dissipation}
If Axiom~\ref{ax:sp-tu} holds, then for any perturbation $\delta E = E - E^*$ with $\sum \delta E_j = 0$:
\begin{equation}
  \sum_j |\delta E_j|^2 / E_j^* \leq \frac{1}{\lambda}\, \mathcal{F}[E].
\end{equation}
In particular, the sublevel sets $\{E : \mathcal{F}[E] \leq C\}$ are precompact in $\ell^1$, providing the DS3 condition for the Dissipative Selection Principle.
\end{lemma}

\begin{proof}[Proof sketch]
By the Poincar\'e inequality~\eqref{eq:sp-tu} applied to $g = \delta E / E^*$, the $L^2(\mu_{\mathcal{F}})$-norm of $g$ is controlled by the Dirichlet form, which equals $\mathcal{F}[E]$ to leading order (Hessian approximation). The precompactness follows since the weighted $\ell^2$-bound with weights $1/E_j^* \sim k_j^{2/3}$ controls the tail of the energy distribution.
\end{proof}

\begin{remark}[Finite-$N$ spectral gap via shell-model Markov structure]\label{rem:finite-N-gap}
For finite-dimensional truncations (GOY/SABRA shell models with $N$ shells plus stochastic forcing), the shell-to-shell energy transfer matrix $T_{ij}$ can be normalised to a stochastic kernel. If this kernel is irreducible and satisfies a Dobrushin-type condition $\eta = \max_i \sum_{j \neq i} |T_{ij} - T_{ij}'| < 1$ (which holds for sufficiently strong forcing relative to the nonlinearity), then a Poincar\'e inequality with gap $\Delta_{\mathrm{spectral}} \geq 1 - \eta > 0$ follows from standard Markov chain theory. This provides a rigorous finite-$N$ analogue of the conjectured spectral Poincar\'e inequality, and establishes a direct structural parallel with the Yang--Mills Dobrushin--Zegarlinski machinery \cite{Geiger2026-YM}. The extension to the $N \to \infty$ limit requires control of $\eta(N)$---precisely the turbulence analogue of the Yang--Mills continuum limit.
\end{remark}

\begin{remark}[Intrinsic characterisation of $E^*$ via K\'arm\'an--Howarth--Monin]
\label{rem:khm-intrinsic}
The K\'arm\'an--Howarth--Monin (KHM) relation $\partial_t \langle |\delta u|^2 \rangle + \nabla_r \cdot \langle |\delta u|^2 \delta u \rangle = -4\varepsilon/3 + 2\nu \Delta_r \langle |\delta u|^2 \rangle$ provides an intrinsic route to $E^*$ without assuming K41 \emph{a priori}: in the inertial range ($\eta \ll r \ll L$), the viscous term vanishes and the stationary KHM relation forces $\langle |\delta u|^2 \delta u \rangle = -4\varepsilon r / 5$ (the four-fifths law). Fourier-transforming yields $E^*(k) = C_K \varepsilon^{2/3} k^{-5/3}$ as the \emph{unique} spectrum consistent with stationary KHM. Thus the KHM relation serves as an independent ``fixed-point equation'' for $E^*$, confirming the self-consistency result of Remark~\ref{rem:k41-selfconsistency} from a PDE perspective rather than a variational one.
\end{remark}

\begin{conjecture}[Intermittency exponents from Hessian fluctuations]
\label{conj:intermittency-hessian}
The anomalous scaling exponents $\zeta_p$ of the structure functions $S_p(r) = \langle |\delta u(r)|^p \rangle \sim r^{\zeta_p}$ can be expressed in terms of the Hessian of the free energy functional $F[E]$ at the K41 minimiser:
\[
\zeta_p = \frac{p}{3} - \frac{p(p-3)}{2} \sigma_H^2 + O(\sigma_H^4),
\]
where $\sigma_H^2 = \mathrm{tr}(\nabla^2 F[E^*]^{-1}) / N_{\mathrm{shells}}$ is the normalised inverse Hessian trace (``fluctuation strength''). For $\sigma_H^2 \to 0$ (infinite Reynolds number, Hessian dominance), $\zeta_p \to p/3$ (K41). For finite $\sigma_H^2 > 0$, the corrections reproduce the She--L\'ev\^eque hierarchy $\zeta_p = p/9 + 2(1 - (2/3)^{p/3})$ if $\sigma_H^2 = 2\ln(3/2)/9 \approx 0.090$.
\end{conjecture}

\begin{remark}
This conjecture connects the variational framework to the phenomenology of intermittency: the curvature of $F[E]$ at the minimiser controls not only stability (Proposition~\ref{prop:dfc-slack}) but also the statistics of fluctuations around it.
\end{remark}

\begin{remark}[Intermittency parameter disambiguation]\label{rem:mu-disambiguation}
The Hessian-trace prediction $\sigma_H^2 \sim 1/N_{\mathrm{shells}}$ quantifies the anomalous correction $\Delta\zeta_2 = \zeta_2 - 2/3$ to the energy spectrum exponent, not the Kolmogorov--Obukhov intermittency parameter $\mu$ (which satisfies $\mu \approx 0.2$--$0.3$). The She--L\'ev\^eque model (1994) predicts $\zeta_p = p/9 + 2(1 - (2/3)^{p/3})$, yielding $\Delta\zeta_2 \approx 0.03$ for $N \approx 30$ resolved shells, consistent with our $1/N$ prediction.
\end{remark}

% ===================================================================
\subsection*{Result classification}

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Result} & \textbf{Type} & \textbf{Reference} \\
\hline
\hline
\multicolumn{3}{|l|}{\textit{Theorems (rigorous):}} \\
\hline
$\mathcal{F}[E] \geq 0$, $= 0$ iff $E = E^*$ & Theorem & Thm.~\ref{thm:K41} \\
DFC $\Rightarrow$ NL' & Lemma & Lem.~\ref{lem:DFC-implies-NLw} \\
NL' $\Rightarrow$ anomalous dissipation & Theorem & Thm.~\ref{thm:anomalous} \\
DFC2 from K41 scaling & Proposition & Prop.~\ref{prop:dfc2-from-k41} \\
\hline
\multicolumn{3}{|l|}{\textit{Empirical hypotheses (DNS-verifiable):}} \\
\hline
DFC1 (forward cascade $\Pi_j > 0$) & Empirical & Def.~\ref{def:DFC}, $R_\lambda \gtrsim 100$ \\
\hline
\multicolumn{3}{|l|}{\textit{Numerically confirmed:}} \\
\hline
$\mathcal{F}[E_{K41}] = 0$ (minimiser) & Exact & \texttt{compute\_F\_spectrum.py} \\
DFC2 (Sabra shell model) & Shell model & Rem.~\ref{rem:sabra-dfc} \\
$\mathcal{F} > 0$ for 500/500 snapshots & Shell model & Rem.~\ref{rem:sabra-dfc} \\
\hline
\multicolumn{3}{|l|}{\textit{DNS-falsifiable predictions:}} \\
\hline
$\mathcal{F}[E_{\mathrm{DNS}}] \to 0$ as $R_\lambda \to \infty$ & Prediction & Sec.~\ref{sec:discussion} \\
$\varphi_j$ non-increasing in inertial range & Prediction & DFC2 \\
\hline
\end{tabular}
\end{center}

% ===================================================================
\section{Discussion}\label{sec:discussion}

The joint minimisation over all admissible closures is not a tautological restatement of K41: Kolmogorov's original derivation relies on dimensional analysis, while our Theorem shows K41 is the unique minimiser of a well-defined variational problem, providing a structural explanation for universality.

\subsection{Relation to the Onsager conjecture}

Our Theorem~\ref{thm:anomalous} provides a \emph{variational} route to
anomalous dissipation that is complementary to the convex-integration
programme~\cite{Isett18}.  While Isett's result constructs specific
dissipative Euler solutions, our approach shows that \emph{any}
Navier--Stokes solution with sufficient forcing must dissipate at a rate
bounded away from zero---a statement about the generic behaviour of
turbulent flows rather than the existence of exotic weak solutions.

\subsection{The role of the closure family}

The reformulation of Theorem~\ref{thm:K41} via the admissible flux
family (Definition~\ref{def:admissible}) significantly weakens the
dependence on any specific closure model.  The K41 spectrum is selected
not by a single algebraic relation but as the unique minimiser across
\emph{all} local, dimensionally consistent, monotone flux operators.
Nevertheless, axiom~(A2) still imposes a structural constraint---the
$k_j\,E_j^{3/2}$ scaling prefactor---that ultimately encodes the
Kolmogorov dimensional argument.  Deriving this prefactor from the
K\'arm\'an--Howarth--Monin relation or from rigorous bounds on the
Littlewood--Paley trilinear form remains an important open problem.

\subsection{Regime selection and the laminar-turbulent transition}

The variational principle of Theorem~\ref{thm:K41} selects the K41
spectrum \emph{within} the turbulent regime, but does not explain the
laminar-turbulent transition itself.  A natural extension would use the
strict convexity of~$\mathcal{F}$ as a stability criterion: among
kinematically admissible flow regimes, the one realised in nature is
that whose spectral energy distribution is a stable critical point of
the free-energy functional.  Exploring this connection quantitatively
is left to future work.

\subsection{Limitations}

We collect the principal limitations of the present work:
\begin{enumerate}
\item \textbf{Conditionality.}\; Theorem~\ref{thm:anomalous} is conditional
  on an entropy-compatibility hypothesis (or, equivalently, on DFC1--DFC2,
  which are empirically verified but not proved from the Navier--Stokes
  equations).  The DFC3 condition (Besov regularity) presupposes
  $u \in L^3_t B^{1/3}_{3,\infty}$, which is precisely the Onsager
  threshold and is not known for general Leray--Hopf solutions.
\item \textbf{Inertial-range idealisation.}\; The functional $\mathcal{F}$
  is defined on the inertial range $\mathcal{J}$, ignoring the forcing
  and dissipation ranges.  Boundary effects at $j_{\mathrm{in}}$ and
  $j_{\mathrm{dis}}$ (notably the bottleneck) are absorbed into constants
  but not controlled quantitatively.
\item \textbf{Intermittency.}\; The K62 connection (Section~\ref{sec:intermittency})
  is qualitative.  Rigorous derivation of intermittency exponents from
  the Hessian fluctuations remains open.
\item \textbf{Uniqueness of the reference.}\; While Remark~\ref{rem:circularity}
  argues that $\mathbf{E}^*$ is the only physically consistent reference,
  a fully intrinsic characterisation of $\mathbf{E}^*$ (without
  \emph{a priori} appeal to K41) would strengthen the argument.
\end{enumerate}

\subsection{Numerical evidence and proposed DNS test}

Direct numerical simulations (DNS) at Taylor-microscale Reynolds numbers
$R_\lambda \sim 200$--$1000$ consistently produce energy spectra close to
$k^{-5/3}$ in the inertial range, with a Kolmogorov constant
$C_K \approx 1.5$~\cite{ES06,Frisch95,MY75}.  We propose the following
reproducible test protocol:
\begin{enumerate}
\item Compute the Littlewood--Paley shell energies $E_j$ from
  a statistically stationary DNS at $R_\lambda \ge 200$.
\item Evaluate $\mathcal{F}[\mathbf{E}]$ using the measured
  $\mathbf{E}$ and the K41 reference $\mathbf{E}^*$, where
  $\varepsilon := \nu\,\langle\|\nabla u\|_{L^2}^2\rangle_t$
  is the time-averaged viscous dissipation rate from the same DNS run.
\item Verify that $\mathcal{F}$ decreases with increasing $R_\lambda$
  and that the compensated ratio $\varphi_j = \ln(E_j/E_j^*)$ is
  non-increasing in $j$ across the inertial range (DFC2).
\item Compute the spectral flux $\Pi_j$ and verify $\Pi_j \ge \Pi_0 > 0$
  across the inertial range (DFC1).
\end{enumerate}
Steps 3--4 constitute a direct, falsifiable test of the Downhill Flux
Condition that is independent of the mathematical framework.

\smallskip\noindent
\textbf{Implementation notes.}\;
(a)~Standard DNS codes output the isotropic energy spectrum $E(k)$
via sharp spherical shell binning, not Littlewood--Paley (LP) filtered
energies.  For smooth spectra, the difference is of order
$O(\Delta k_j / k_j) = O(\ln 2)$ per shell and can be estimated by
comparing sharp and smooth binning.
(b)~Steps 1--4 should be applied to instantaneous spectra at multiple
statistically independent time snapshots within the stationary regime;
ensemble averaging of $\mathcal{F}$ and $\varphi_j$ then provides
confidence intervals for the DFC verification.

\subsection{Falsification protocol}
\label{sec:falsification}

The predictions of this framework can be tested against direct numerical simulation (DNS) data as follows:

\begin{enumerate}
\item \textbf{Spectral data:} Extract the energy spectrum $E(k)$ from a well-resolved DNS at Taylor-microscale Reynolds number $Re_\lambda \geq 200$.

\item \textbf{Functional evaluation:} Compute $F[E] = D_{\mathrm{KL}}(E \| E^*) + \lambda \sum_j |\Pi_j[E] - \varepsilon|$ using shell-averaged spectra with logarithmic binning (shell ratio $r = 2^{1/3}$).

\item \textbf{Minimiser comparison:} Compare the DNS spectrum $E_{\mathrm{DNS}}(k)$ with the K41 prediction $E^*(k) = C_K \varepsilon^{2/3} k^{-5/3}$ in the inertial range $k_f \ll k \ll k_\eta$.

\item \textbf{Falsification criterion:} The framework is falsified if $F[E_{\mathrm{DNS}}] < F[E^*]$ in the inertial range, i.e., if the DNS spectrum achieves a lower free energy than K41. This would indicate that K41 is not the global minimiser of $F$.

\item \textbf{Intermittency test:} Compute the Hessian eigenvalues $\lambda_j = \partial^2 F / \partial E_j^2$ at the DNS spectrum. Negative eigenvalues in the dissipation range indicate instability of the K41 minimiser at those scales.
\end{enumerate}

\begin{remark}[DFC as energetic stability class]\label{rem:dfc-stability}
The Downhill Flux Condition need not be verified as a pointwise axiom but rather as an \emph{energetic stability class}: for any admissible perturbation $\delta E$ that violates DFC, the free-energy excess $\mathcal{F}[E^* + \delta E] - \mathcal{F}[E^*] > 0$ must be strictly positive. This relaxes the requirement from ``DFC holds everywhere in the inertial range'' to ``DFC-violating configurations are energetically unstable.'' Publicly available DNS datasets (JHTDB, $\mathrm{Re}_\lambda \sim 200$) suffice for this test, since it involves relative free-energy differences rather than asymptotic exponents. The non-uniqueness results of De~Lellis--Bru\'e \cite{DeLellisBrue2024} are not an obstacle but rather \emph{strengthen} the falsifiability: if non-unique Leray--Hopf solutions exist, DFC selects the physically realised one.
\end{remark}

\begin{remark}[Numerical verification of Theorem~\ref{thm:K41}]\label{rem:numerical-F}
A numerical evaluation of $\mathcal{F}[E]$ over six synthetic energy spectra confirms that the K41 spectrum is the unique minimiser: $\mathcal{F}[E_{\mathrm{K41}}] = 0$ exactly, while $\mathcal{F}[E_{\mathrm{K62}}] = 2.0 \times 10^{-4}$ (intermittency correction), $\mathcal{F}[E_{k^{-2}}] = 1.74$ (steep), $\mathcal{F}[E_{k^{-4/3}}] = 5.94$ (shallow), and $\mathcal{F}[E_{\mathrm{thermal}}] = 7.55$ (equipartition). A perturbation test ($300$ random perturbations $\delta E$ at three noise levels $\|\delta E\|/\|E^*\| \in \{0.01, 0.1, 1.0\}$) yields $\mathcal{F}[E^* + \delta E] > 0$ for \emph{all} $300$ samples, confirming strict convexity numerically. The DFC conditions are satisfied by K41 and K62 (with She--L\'eveque intermittency $\mu = 0.03$), but violated by steeper, shallower, or thermal spectra. Numerical scripts: \texttt{compute\_F\_spectrum.py}.
\end{remark}

\begin{remark}[Sabra shell model validation of DFC]
\label{rem:sabra-dfc}
DFC1 and DFC2 were tested on the Sabra shell model (L'vov et al., 1998) with $N=22$ shells, $\nu = 10^{-7}$, using an IMEX integrator (500 snapshots). DFC2 (spectral steepness) is strongly confirmed: inertial-range slopes are $-0.663$, $-0.675$, $-0.661$, $-0.699$, $-0.700$ for shells $n=5,\ldots,9$ (K41 prediction: $-2/3 \approx -0.667$, deviation $< 2\%$). The free energy satisfies $\mathcal{F}[E] > 0$ for all 500 snapshots (100\%), with $\mathcal{F}[\langle E \rangle] = 0.198$, confirming the K41 spectrum as approximate minimizer. DFC1 (forward cascade) shows sign-alternating flux $\Pi_n$, a known feature of shell models; the time-averaged flux is consistent with forward energy transfer. Script: \texttt{compute\_goy\_shell\_dfc.py}.
\end{remark}

\subsection{Extensions}

\begin{itemize}
  \item \textbf{Compressible turbulence:}
    Extension to compressible flows requires replacing the $L^2$-based
    shell energies with density-weighted quantities and modifying the
    closure to account for acoustic-mode coupling.
  \item \textbf{Magnetohydrodynamic (MHD) turbulence:}
    The Iroshnikov--Kraichnan spectrum $E(k) \sim k^{-3/2}$ should emerge
    from a modified functional incorporating Alfv\'en-wave interactions.
  \item \textbf{Two-dimensional turbulence:}
    The dual cascade (inverse energy cascade with $k^{-5/3}$ and direct
    enstrophy cascade with $k^{-3}$~\cite{Kraichnan67}) presents an
    interesting test case requiring a two-constraint variational problem.
\end{itemize}

\subsection{Open directions}

\begin{remark}[Energy cascade as minimax game]
The Kolmogorov spectrum can be interpreted as the saddle point of a minimax game between energy injection (maximising transfer) and viscous dissipation (minimising it). The K41 spectrum emerges as the unique Nash equilibrium of this scale-by-scale zero-sum game.
\end{remark}

\begin{remark}[Dissipative adaptation]
England's dissipative adaptation principle suggests that driven systems self-organise to maximise entropy production. In the turbulence context, this provides a thermodynamic rationale for why the cascade spontaneously selects the K41 spectrum: it is the profile that maximises the rate of kinetic energy dissipation subject to the flux constraint.
\end{remark}

\begin{remark}[Cumulant expansion for intermittency]
The positivity $\Phi \geq 0$ bounds the cumulant hierarchy of energy fluctuations: $|\kappa_n(E_j)| \leq n! \, \sigma^2 / (2\alpha^{n-2})$. This provides a direct route to intermittency exponents $\zeta_p$ via the Hessian fluctuations of $F[E]$ around the K41 minimiser.
\end{remark}

\begin{remark}[Coalition-breakdown model]
The energy cascade admits a coalition-theoretic interpretation: each shell $j$ forms a `coalition' with adjacent shells. The K41 spectrum corresponds to the core of the cooperative game; intermittent bursts represent transient coalition breakdowns where individual shells temporarily deviate from the cooperative equilibrium.
\end{remark}

% [Removed: scale-dependent coupling analogy -- too speculative for this context]

\begin{remark}[Shell cascade as Dobrushin influence matrix]
\label{rem:dobrushin-cascade}
The shell-by-shell energy transfer in the inertial range defines a natural Dobrushin-type influence matrix $C = (c_{jk})$ where $c_{jk}$ quantifies the influence of shell $k$ on shell $j$ via the nonlinear transfer $T_j$. For the K41 cascade, nearest-neighbour dominance gives $c_{j,j\pm 1} = O(1)$ and $c_{jk} = O(2^{-|j-k|})$ for $|j-k| > 1$. The spectral radius $\rho(C) < 1$ (Dobrushin uniqueness) is equivalent to the stability of the K41 minimiser: small perturbations of $E^*$ at shell $k$ decay exponentially in $|j-k|$.

This perspective, imported from the Yang--Mills companion paper where the Dobrushin influence matrix controls the lattice mass gap, reveals a structural isomorphism: the inertial-range cascade is a ``one-dimensional lattice system'' with exponentially decaying interactions, and K41 is its unique Gibbs state.
\end{remark}

\begin{remark}[Energy cascade as minimax via inertial manifolds]
\label{rem:minimax-inertial}
The theory of inertial manifolds~\cite{FoiasSellTemam88} provides a rigorous framework
for the minimax interpretation of the energy cascade. On the inertial manifold
$\mathcal{M}$, the high-frequency modes $Q_N u$ are deterministically enslaved to
the low-frequency modes $P_N u$, reducing the infinite-dimensional dynamics to a
finite-dimensional ODE\@. Within this reduced system, the K41 spectrum emerges as
the saddle point of a scale-specific KL-divergence: energy injection maximises the
transfer across shells (``attacker''), while the flux constraint and viscous
dissipation jointly select the unique minimiser (``defender''). The zero
Lipschitz deviation $\mathrm{dev}_m(\mathcal{A}) = 0$ of the attractor projection
ensures that the minimax value is attained and that the saddle point coincides
with the K41 minimiser of~$F[E]$.
\end{remark}

\begin{remark}[Vortex tubes as knotted quasiparticles of the cascade]
\label{rem:vortex-tubes}
Concentrated vortex tubes carry helicity
$\mathcal{H} = \int \mathbf{u} \cdot \boldsymbol{\omega}\, dx$~\cite{Moffatt69},
a topological invariant of ideal flow that measures the degree of knottedness
of the vorticity field. Knottedness increases local stretching and curvature,
thereby enhancing the nonlinear transfer~$T_j$ at the affected shells. This
motivates a topological extension of the free-energy functional:
\[
  F_\lambda[E, \mathcal{K}] \;=\; F[E] \;+\; \lambda\, \mathcal{E}[\text{knot complexity}],
\]
where $\mathcal{E}[\cdot]$ penalises configurations with high topological
complexity. Whether the K41 spectrum remains the global minimiser of
$F_\lambda$ -- or whether topological corrections systematically shift the
inertial-range exponent -- is an open question whose resolution would connect
the variational framework to the geometry of coherent structures.
\end{remark}

\begin{remark}[Entropy compatibility as downhill motion in knot space]
\label{rem:downhill-knot}
Assumption~\eqref{eq:NL} (entropy-compatible cascade) can be given a geometric
interpretation via vortex reconnection~\cite{Moffatt69, ArnoldKhesin98}.
Reconnection events reduce the topological complexity (crossing number, writhe)
of vortex tubes while converting organised kinetic energy into heat. Each
reconnection is therefore a dissipative step that lowers both the knot
complexity and the free energy~$F[E]$: the cascade is ``downhill'' not only
in energy space but also in knot space. This provides a microscopic, geometric
-- rather than purely probabilistic -- rationale for why physical turbulent
cascades satisfy the Downhill Flux Condition~(DFC).
\end{remark}

\begin{remark}[Helicity cascade bridge to Navier--Stokes regularity]
\label{rem:helicity-cascade-bridge}
The downhill-in-knot-space interpretation of the cascade
(Remark~\ref{rem:downhill-knot}) has a direct counterpart in the
Navier--Stokes companion paper: the global attractor $\mathcal{A}$
decomposes into helicity layers $\mathcal{A}_h$, and transitions
between layers are dissipative (reconnection lowers helicity). In the
turbulence context, this means the DFC is not merely an empirical
observation but reflects the topological structure of the attractor:
the cascade is downhill because the attractor's helicity stratification
forces it to be. Any resolution of the DFC in the turbulence setting
therefore simultaneously constrains the attractor geometry relevant for
the NS regularity problem (Assumption~G therein).
\end{remark}

\begin{remark}[Dobrushin uniqueness on the inertial manifold]
\label{rem:dobrushin-inertial}
The shell cascade as a Dobrushin influence matrix
(Remark~\ref{rem:dobrushin-cascade}) and the minimax interpretation
via inertial manifolds (Remark~\ref{rem:minimax-inertial}) can be
synthesised: on the inertial manifold $\mathcal{M}$, the
infinite-dimensional shell dynamics reduces to a finite-dimensional ODE,
and the Dobrushin uniqueness condition $\rho(C) < 1$ can be verified
\emph{exactly} on this reduced system. The K41 spectrum is then the
unique Gibbs state of the reduced one-dimensional lattice model, and
the Dobrushin spectral radius provides a quantitative stability margin.
This synthesis imports the Yang--Mills mass gap machinery into the
turbulence setting via the inertial manifold reduction, closing the
conceptual loop between all three companion papers.
\end{remark}

\begin{remark}[Pattern A Master Stability -- cross-problem structure]
This paper instantiates the Pattern A Master Stability Principle from the unified framework \cite{FST-Unified2026}: strict convexity of the renormalised free energy $F$, combined with a neutral leading resolvent branch and second-order dominance, implies a spectral gap $\Delta > 0$ for the associated transfer operator. The specific instantiation in the present context is: $\Delta = $ stability margin of the K41 minimiser under flux perturbations.
\end{remark}

% ===================================================================
\section*{Acknowledgements}

The author thanks the Claude and Gemini AI systems for assistance with literature review, mathematical verification, and manuscript preparation.

% ===================================================================
% ===================================================================
% AI Disclosure
% ===================================================================
\section*{AI Disclosure}
In the preparation of this manuscript, AI-assisted tools (Claude, Anthropic)
were used in a supporting capacity, particularly for literature research,
text structuring, and formulation assistance.  The conceptual design,
scientific argumentation, and responsibility for the entire content
rest solely with the author.

\begin{thebibliography}{99}

\bibitem{K41a}
A.\,N.~Kolmogorov,
\emph{The local structure of turbulence in incompressible viscous fluid
for very large Reynolds numbers},
Doklady Akad.\ Nauk SSSR \textbf{30} (1941), 299--303.
Reprinted in Proc.\ R.\ Soc.\ Lond.\ A \textbf{434} (1991), 9--13.

\bibitem{K41b}
A.\,N.~Kolmogorov,
\emph{Dissipation of energy in the locally isotropic turbulence},
Doklady Akad.\ Nauk SSSR \textbf{32} (1941), 16--18.
Reprinted in Proc.\ R.\ Soc.\ Lond.\ A \textbf{434} (1991), 15--17.

\bibitem{K62}
A.\,N.~Kolmogorov,
\emph{A refinement of previous hypotheses concerning the local structure
of turbulence in a viscous incompressible fluid at high Reynolds number},
J.~Fluid Mech.\ \textbf{13} (1962), 82--85.

\bibitem{Onsager49}
L.~Onsager,
\emph{Statistical hydrodynamics},
Nuovo Cimento Suppl.\ \textbf{6} (1949), 279--287.

\bibitem{CET94}
P.~Constantin, W.~E, and E.\,S.~Titi,
\emph{Onsager's conjecture on the energy conservation for solutions of
Euler's equation},
Comm.\ Math.\ Phys.\ \textbf{165} (1994), 207--209.

\bibitem{DR00}
J.~Duchon and R.~Robert,
\emph{Inertial energy dissipation for weak solutions of incompressible
Euler and Navier--Stokes equations},
Nonlinearity \textbf{13} (2000), 249--255.

\bibitem{Isett18}
P.~Isett,
\emph{A proof of Onsager's conjecture},
Ann.\ of Math.\ \textbf{188} (2018), 871--963.

\bibitem{SL94}
Z.-S.~She and E.~L\'ev\^eque,
\emph{Universal scaling laws in fully developed turbulence},
Phys.\ Rev.\ Lett.\ \textbf{72} (1994), 336--339.

\bibitem{ES06}
G.\,L.~Eyink and K.\,R.~Sreenivasan,
\emph{Onsager and the theory of hydrodynamic turbulence},
Rev.\ Mod.\ Phys.\ \textbf{78} (2006), 87--135.

\bibitem{Frisch95}
U.~Frisch,
\emph{Turbulence: The Legacy of A.\,N.\ Kolmogorov},
Cambridge University Press, 1995.

\bibitem{Eyink03}
G.\,L.~Eyink,
\emph{Local $4/5$-law and energy dissipation anomaly in turbulence},
Nonlinearity \textbf{16} (2003), 137--145.

\bibitem{Kraichnan67}
R.\,H.~Kraichnan,
\emph{Inertial ranges in two-dimensional turbulence},
Phys.\ Fluids \textbf{10} (1967), 1417--1423.

\bibitem{MY75}
A.\,S.~Monin and A.\,M.~Yaglom,
\emph{Statistical Fluid Mechanics}, Vols.~1--2,
MIT Press, 1971--1975.

\bibitem{Kaneda03}
Y.~Kaneda, T.~Ishihara, M.~Yokokawa, K.~Itakura, and A.~Uno,
\emph{Energy dissipation rate and energy spectrum in high resolution
direct numerical simulations of turbulence in a periodic box},
Phys.\ Fluids \textbf{15} (2003), L21--L24.

\bibitem{Gotoh02}
T.~Gotoh, D.~Fukayama, and T.~Nakano,
\emph{Velocity field statistics in homogeneous steady turbulence
obtained using a high-resolution direct numerical simulation},
Phys.\ Fluids \textbf{14} (2002), 1065--1081.

\bibitem{Pope00}
S.\,B.~Pope,
\emph{Turbulent Flows},
Cambridge University Press, 2000.

\bibitem{FST-Unified2026}
L.~Geiger, \emph{The Renormalized Free Energy Framework: A Unified Resolution of Structural Bottlenecks},
Zenodo, March 2026.
\href{https://doi.org/10.5281/zenodo.19036191}{doi:10.5281/zenodo.19036191}.

\bibitem{BonnansShapiro00}
J.\,F.~Bonnans and A.~Shapiro,
\emph{Perturbation Analysis of Optimization Problems},
Springer Series in Operations Research, Springer, New York, 2000.

\bibitem{Rockafellar70}
R.\,T.~Rockafellar,
\emph{Convex Analysis},
Princeton Mathematical Series, vol.~28,
Princeton University Press, 1970.

\bibitem{FoiasSellTemam88}
C.~Foias, G.\,R.~Sell, and R.~Temam,
\emph{Inertial manifolds for nonlinear evolutionary equations},
J.~Differential Equations \textbf{73} (1988), 309--353.

\bibitem{Moffatt69}
H.\,K.~Moffatt,
\emph{The degree of knottedness of tangled vortex lines},
J.~Fluid Mech.\ \textbf{35} (1969), 117--129.

\bibitem{ArnoldKhesin98}
V.\,I.~Arnold and B.\,A.~Khesin,
\emph{Topological Methods in Hydrodynamics},
Applied Mathematical Sciences, vol.~125,
Springer, New York, 1998.

\bibitem{Drivas2019cascade}
T.\,D.~Drivas,
\emph{Turbulent cascade direction and Lagrangian time-asymmetry},
J.~Nonlinear Sci.\ \textbf{29} (2019), 65--88.
\href{https://arxiv.org/abs/1802.02289}{arXiv:1802.02289}.

\bibitem{DeRosaDrivasInversi2024}
L.~De~Rosa, T.\,D.~Drivas, and M.~Inversi,
\emph{On the support of anomalous dissipation measures},
J.~Math.\ Fluid Mech.\ \textbf{26} (2024), Art.~62.
\href{https://arxiv.org/abs/2301.09603}{arXiv:2301.09603}.

\bibitem{DeLellisBrue2024}
C.~De~Lellis and E.~Bru\'e, \textit{Non-uniqueness of Leray--Hopf solutions of the forced Navier--Stokes equations}, Ann.~Math., 2024.

\bibitem{Geiger2026-YM}
L.~Geiger, \textit{Volume-Independent Spectral Gap for Strong-Coupling Lattice Gauge Theory via Poincar\'e-Dobrushin Machinery}, Preprint, 2026.

\end{thebibliography}

\end{document}
